---
title: "nationalrat vote share histogram"
output: html_document
date: "2026-06-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


##Nationalrat Vote Share Histogram
Plot the vote share values of the top 6 parties over the years 1971 to 2023 in a histogram to see what vote share values are most common. This will later be compared to the vote shares simulated by STAN in the prior predictive check to see how the values of the coefficients need to be adjusted for the forecasting model to later work with real data.

```{r}
library(rstan)
library(dplyr)
library(ggplot2)
library(scales)
library(tidyr)
library(readr)

```


```{r}
#party vote shares 1971-2023

party_levels <- c("fdp", "cvp", "sp", "svp", "other")
predictor_cols <- c("smi_norm", "health_insurance_norm", "immigration_norm")

party_colors <- c(
  "fdp" = "cornflowerblue",
  "cvp" = "darkorange",
  "sp" = "red2",
  "svp" = "green4",
  "other" = "plum2"
)

election_data <- read.csv("swiss election data.csv")

election_data <- election_data %>%
  mutate(
    party = factor(party, levels = party_levels),
    year = as.integer(year)
  ) %>%
  arrange(year, party)

#fdp
fdp_results <- election_data %>%
  filter(party == "fdp")

fdp <- c(fdp_results$vote_share)

#cvp
cvp_results <- election_data %>%
  filter(party == "cvp")

cvp <- c(cvp_results$vote_share)

#sp
sp_results <- election_data %>%
  filter(party == "sp")

sp <- c(sp_results$vote_share)

#svp
svp_results <- election_data %>%
  filter(party == "svp")

svp <- c(svp_results$vote_share)

#other
other_results <- election_data %>%
  filter(party == "other")

other <- c(other_results$vote_share)

```


#Stats by Party
```{r}
#FDP
fdp_min <- min(fdp)
fdp_max <- max(fdp)
fdp_mean <- mean(fdp)
fdp_median <- median(fdp)

bin_width <- 0.01
#create breaks explicitly
fdp_breaks <- seq(0, max(fdp) + bin_width, by = bin_width)
#cut data into bins
fdp_bins <- cut(fdp, breaks = fdp_breaks, include.lowest = TRUE, right = TRUE)
#count frequencies
fdp_bin_counts <- table(fdp_bins)
#find index of the highest bin
fdp_highest_bin_index <- which.max(fdp_bin_counts)
#calculate center directly from breaks vector
fdp_lower_bound <- fdp_breaks[fdp_highest_bin_index]
fdp_upper_bound <- fdp_breaks[fdp_highest_bin_index + 1]
fdp_modal_class_center <- (fdp_lower_bound + fdp_upper_bound) / 2

cat("FDP", "\nMin: ", fdp_min, "\nMax: ", fdp_max, "\nMean: ", fdp_mean, "\nMedian: ", fdp_median, "\nMode: ", fdp_modal_class_center)
```


```{r}
#CVP
cvp_min <- min(cvp)
cvp_max <- max(cvp)
cvp_mean <- mean(cvp)
cvp_median <- median(cvp)

bin_width <- 0.01
#create breaks explicitly
cvp_breaks <- seq(0, max(cvp) + bin_width, by = bin_width)
#cut data into bins
cvp_bins <- cut(cvp, breaks = cvp_breaks, include.lowest = TRUE, right = TRUE)
#count frequencies
cvp_bin_counts <- table(cvp_bins)
#find index of the highest bin
cvp_highest_bin_index <- which.max(cvp_bin_counts)
#calculate center directly from breaks vector
cvp_lower_bound <- cvp_breaks[cvp_highest_bin_index]
cvp_upper_bound <- cvp_breaks[cvp_highest_bin_index + 1]
cvp_modal_class_center <- (cvp_lower_bound + cvp_upper_bound) / 2

cat("CVP", "\nMin: ", cvp_min, "\nMax: ", cvp_max, "\nMean: ", cvp_mean, "\nMedian: ", cvp_median, "\nMode: ", cvp_modal_class_center)

```


```{r}
#SP
sp_min <- min(sp)
sp_max <- max(sp)
sp_mean <- mean(sp)
sp_median <- median(sp)

bin_width <- 0.01
#create breaks explicitly
sp_breaks <- seq(0, max(sp) + bin_width, by = bin_width)
#cut data into bins
sp_bins <- cut(sp, breaks = sp_breaks, include.lowest = TRUE, right = TRUE)
#count frequencies
sp_bin_counts <- table(sp_bins)
#find index of the highest bin
sp_highest_bin_index <- which.max(sp_bin_counts)
#calculate center directly from breaks vector
sp_lower_bound <- sp_breaks[sp_highest_bin_index]
sp_upper_bound <- sp_breaks[sp_highest_bin_index + 1]
sp_modal_class_center <- (sp_lower_bound + sp_upper_bound) / 2

cat("SP", "\nMin: ", sp_min, "\nMax: ", sp_max, "\nMean: ", sp_mean, "\nMedian: ", sp_median, "\nMode: ", sp_modal_class_center)

```


```{r}
#SVP
svp_min <- min(svp)
svp_max <- max(svp)
svp_mean <- mean(svp)
svp_median <- median(svp)

bin_width <- 0.01
#create breaks explicitly
svp_breaks <- seq(0, max(svp) + bin_width, by = bin_width)
#cut data into bins
svp_bins <- cut(svp, breaks = svp_breaks, include.lowest = TRUE, right = TRUE)
#count frequencies
svp_bin_counts <- table(svp_bins)
#find index of the highest bin
svp_highest_bin_index <- which.max(svp_bin_counts)
#calculate center directly from breaks vector
svp_lower_bound <- svp_breaks[svp_highest_bin_index]
svp_upper_bound <- svp_breaks[svp_highest_bin_index + 1]
svp_modal_class_center <- (svp_lower_bound + svp_upper_bound) / 2

cat("SVP", "\nMin: ", svp_min, "\nMax: ", svp_max, "\nMean: ", svp_mean, "\nMedian: ", svp_median, "\nMode: ", svp_modal_class_center)

```


```
#Grüne
green_min <- min(green)
green_max <- max(green)
green_mean <- mean(green)
green_median <- median(green)

bin_width <- 0.01
#create breaks explicitly
green_breaks <- seq(0, max(green) + bin_width, by = bin_width)
#cut data into bins
green_bins <- cut(green, breaks = green_breaks, include.lowest = TRUE, right = TRUE)
#count frequencies
green_bin_counts <- table(green_bins)
#find index of the highest bin
green_highest_bin_index <- which.max(green_bin_counts)
#calculate center directly from breaks vector
green_lower_bound <- breaks[green_highest_bin_index]
green_upper_bound <- breaks[green_highest_bin_index + 1]
green_modal_class_center <- (green_lower_bound + green_upper_bound) / 2

cat("Grüne", "\nMin: ", green_min, "\nMax: ", green_max, "\nMean: ", green_mean, "\nMedian: ", green_median, "\nMode: ", green_modal_class_center)

```


```{r}
#OTHER
other_min <- min(other)
other_max <- max(other)
other_mean <- mean(other)
other_median <- median(other)

bin_width <- 0.01
#create breaks explicitly
other_breaks <- seq(0, max(other) + bin_width, by = bin_width)
#cut data into bins
other_bins <- cut(other, breaks = other_breaks, include.lowest = TRUE, right = TRUE)
#count frequencies
other_bin_counts <- table(other_bins)
#find index of the highest bin
other_highest_bin_index <- which.max(other_bin_counts)
#calculate center directly from breaks vector
other_lower_bound <- other_breaks[other_highest_bin_index]
other_upper_bound <- other_breaks[other_highest_bin_index + 1]
other_modal_class_center <- (other_lower_bound + other_upper_bound) / 2

cat("Other", "\nMin: ", other_min, "\nMax: ", other_max, "\nMean: ", other_mean, "\nMedian: ", other_median, "\nMode: ", other_modal_class_center)

```


#Plot Total
```{r}
#one list of all vote share values
all_shares <- c(fdp, cvp, sp, svp, other)

#data frame
df <- data.frame(vote_share = all_shares)

```


```{r}
library(ggplot2)
library(dplyr)

#stats
avg_vote_share <- mean(all_shares)
min_vote_share <- min(all_shares)
max_vote_share <- max(all_shares)
median_vote_share <- median(all_shares)

#computing the mode

bin_width <- 0.01
#create breaks explicitly
breaks <- seq(0, max_vote_share + bin_width, by = bin_width)
#cut data into bins
bins <- cut(all_shares, breaks = breaks, include.lowest = TRUE, right = TRUE)
#count frequencies
bin_counts <- table(bins)
#find index of the highest bin
highest_bin_index <- which.max(bin_counts)
#calculate center directly from breaks vector
lower_bound <- breaks[highest_bin_index]
upper_bound <- breaks[highest_bin_index + 1]
modal_class_center <- (lower_bound + upper_bound) / 2


#histogram
p <- ggplot(df, aes(x=vote_share)) +
  geom_histogram(binwidth = 0.01, boundary = 0, color = "black", fill = "lightblue") +
  geom_vline(xintercept = avg_vote_share, color = "red", linewidth = 1, linetype = "solid") +
  geom_vline(xintercept = c(min_vote_share, max_vote_share), color = "darkgreen", linewidth = 1, linetype = "dashed") +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  labs(
    x = "Vote Share",
    y = "Frequency"
  ) +
  theme_minimal() +
  theme(
    margins = margin(t = 10, r = 10, b = 10, l = 10, unit = "pt")
  )

print(p)

ggsave("Vote Share Distribution 1971-2023 (Bin Width 0.01).png", p, width = 6, height = 4, dpi = 200, units = "in")

#summary
cat("Minimum Vote Share: ", min_vote_share,
    "\nMaximum Vote Share: ", max_vote_share,
    "\nAverage Vote Share: ", avg_vote_share,
    "\nMedian Vote Share: ", median_vote_share,
    "\nModal Bin Center: ", modal_class_center)

```

```{r}
hist(all_shares,
     col = "lightblue",
     border = "white",
     breaks = 84,
     xlim = c(0, 0.35),
     main = "Vote Shares")

abline(v = avg_vote_share, col = "red", lwd = 2)
abline(v = median_vote_share, col = "purple", lwd = 2)
abline(v = modal_class_center, col = "orange", lwd = 2)

abline(v = min_vote_share, col = "green", lwd = 2)
abline(v = max_vote_share, col = "green", lwd = 2)
```












