---
title: "run_dynamic_model_1m"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
library(rstan)
library(readr)
library(tidyr)
library(ggplot2)
library(dplyr)
library(scales)
library(bayesplot)

```

###Colors
```{r}
#party colors and preferred order of parties

party_colors <- c(
  "fdp" = "cornflowerblue",
  "cvp" = "darkorange",
  "sp" = "red2",
  "svp" = "green4",
  "other" = "plum2"
)

party_levels <- c("fdp", "cvp", "sp", "svp", "other")
party_labels <- c(
  "fdp" = "FDP",
  "cvp" = "CVP",
  "sp" = "SP",
  "svp" = "SVP",
  "other" = "Other"
)

#predictors
predictor_cols <- c("smi_norm", "health_insurance_norm", "immigration_norm")
```

###Election Data

```{r}
#election data
election_data <- read_csv("swiss election data.csv")

results_2019 <- election_data %>%
  filter(year == 2019) %>%
  mutate(vote_share = vote_share *100)
results_2019$party <- factor(results_2019$party, levels = party_levels)

results_2023 <- election_data %>%
  filter(year == 2023) %>%
  mutate(vote_share = vote_share *100)
results_2023$party <- factor(results_2023$party, levels = party_levels)

```

###Polling Data

Sort through the Swiss polling data.

```{r}
swiss_polling_data <- read_csv("swiss polling data.csv")

company_labels <- c(
  "fors" = "FORS",
  "sotomo" = "Sotomo",
  "tamedia" = "Tamedia"
)

#convert polling_period strings to date
swiss_polling_data$polling_period <- as.Date(swiss_polling_data$polling_period, format = "%d.%m.%Y")

#separate the polling data by election year
SPD_2019 <- swiss_polling_data %>%
  filter(year == 2019)

SPD_2023 <- swiss_polling_data %>%
  filter(year == 2023)

```

```{r}
#election days
T_2019 <- as.Date("2019-10-20")
T_2023 <- as.Date("2023-10-22")

```

```{r}
#list of unique polling days (60 weeks days prior to election day)
start_2019 = T_2019 - 413
start_2023 = T_2023 - 413

all_weeks_2019 <- seq(from = start_2019, to = T_2019, by = "week")
all_weeks_2023 <- seq(from = start_2023, to = T_2023, by = "week")

if (tail(all_weeks_2019, 1) != T_2019) {
  all_weeks_2019 <- c(all_weeks_2019, T_2019)
  message("Added Election Day to 2019 timeline manually.")
}

if (tail(all_weeks_2023, 1) != T_2023) {
  all_weeks_2023 <- c(all_weeks_2023, T_2023)
  message("Added Election Day to 2023 timeline manually")
}
```


```{r}
#data frames with time id and days to election

SPD_2019_time <- SPD_2019 %>%
  mutate(
    days_from_start = as.numeric(polling_period - start_2019),
    time_id = floor(days_from_start / 7) + 1,
    dte = as.numeric(difftime(T_2019, polling_period, units = "days"))
    ) %>%
  filter(time_id >= 1 & time_id <= length(all_weeks_2019))

SPD_2023_time <- SPD_2023 %>%
  mutate(
    days_from_start = as.numeric(polling_period - start_2023),
    time_id = floor(days_from_start / 7) + 1,
    dte = as.numeric(difftime(T_2023, polling_period, units = "days"))
    ) %>%
  filter(time_id >= 1 & time_id <= length(all_weeks_2023))

```

```{r}
#polling data filtered so remove the last 1 Month of data
cutoff_1m_2019 <- T_2019 - 30
cutoff_1m_2023 <- T_2023 - 30

SPD_2019_time <- SPD_2019_time %>%
  filter(polling_period <= cutoff_1m_2019)

SPD_2023_time <- SPD_2023_time %>%
  filter(polling_period <= cutoff_1m_2023)
```


```{r}
#company ids <-  changed order so sotomo is baseline
companies <- c("tamedia", "fors", "sotomo")

SPD_2019_time_comp <- SPD_2019_time %>%
  mutate(
    company_id = match(polling_company, companies)
  )

SPD_2023_time_comp <- SPD_2023_time %>%
  mutate(
    company_id = match(polling_company, companies)
  )

```

```{r}
#poll id
poll_id_2019 <- SPD_2019_time_comp %>%
  distinct(polling_company, time_id) %>%
  arrange(time_id, polling_company) %>%
  mutate(poll_id = row_number())

SPD_2019_id <- SPD_2019_time_comp %>%
  left_join(poll_id_2019, by = c("polling_company", "time_id"))

SPD_2019_id <- SPD_2019_id %>%
  filter(!is.na(time_id)) %>%
  arrange(poll_id)

poll_id_2023 <- SPD_2023_time_comp %>%
  distinct(polling_company, time_id) %>%
  arrange(time_id, polling_company) %>%
  mutate(poll_id = row_number())

SPD_2023_id <- SPD_2023_time_comp %>%
  left_join(poll_id_2023, by = c("polling_company", "time_id"))

SPD_2023_id <- SPD_2023_id %>%
  filter(!is.na(time_id)) %>%
  arrange(poll_id)

```

```{r}
#y for Stan [N_polls, N_p]
y_2019 <- SPD_2019_id %>%
  select(poll_id, party, counts) %>%
  pivot_wider(
    id_cols = poll_id,
    names_from = party,
    values_from = counts,
    values_fill = 0
  ) %>%
  select(all_of(party_levels)) %>%
  as.matrix()

y_2023 <- SPD_2023_id %>%
  select(poll_id, party, counts) %>%
  pivot_wider(
    id_cols = poll_id,
    names_from = party,
    values_from = counts,
    values_fill = 0
  ) %>%
  select(all_of(party_levels)) %>%
  as.matrix()

```

```{r}
poll_metadata_2019 <- SPD_2019_id %>%
  group_by(poll_id) %>%
  slice(1) %>%
  ungroup()

poll_metadata_2023 <- SPD_2023_id %>%
  group_by(poll_id) %>%
  slice(1) %>%
  ungroup()

```

Import EWMA csv files
```{r}
ewma_2019 <- read_csv("ewma_2019.csv")
ewma_2019_1m <- ewma_2019 %>%
  filter(polling_period >= start_2019 & polling_period<= cutoff_1m_2019)

ewma_2023 <- read_csv("ewma_2023.csv")
ewma_2023_1m <- ewma_2023 %>%
  filter(polling_period >= start_2023 & polling_period<= cutoff_1m_2023)

```


###Fundamentals Data

Sort through the Swiss election data.

```{r}
fundamentals_2019 <- read_csv("fundamentals_predictions_2019.csv")
vE_2019 <- fundamentals_2019$mean_pred[match(party_levels, fundamentals_2019$party)]
vE_sd_2019 <- fundamentals_2019$sd_pred[match(party_levels, fundamentals_2019$party)]

```

```{r}
fdp_pred_2019 <- fundamentals_2019 %>%
  filter(party == "fdp")

cvp_pred_2019 <- fundamentals_2019 %>%
  filter(party == "cvp")

sp_pred_2019 <- fundamentals_2019 %>%
  filter(party == "sp")

svp_pred_2019 <- fundamentals_2019 %>%
  filter(party == "svp")

other_pred_2019 <- fundamentals_2019 %>%
  filter(party == "other")

```


```{r}
fundamentals_2023 <- read_csv("fundamentals_predictions_2023.csv")

vE_2023 <- fundamentals_2023$mean_pred[match(party_levels, fundamentals_2023$party)]
vE_sd_2023 <- fundamentals_2023$sd_pred[match(party_levels, fundamentals_2023$party)]

```

```{r}
fdp_pred_2023 <- fundamentals_2023 %>%
  filter(party == "fdp")

cvp_pred_2023 <- fundamentals_2023 %>%
  filter(party == "cvp")

sp_pred_2023 <- fundamentals_2023 %>%
  filter(party == "sp")

svp_pred_2023 <- fundamentals_2023 %>%
  filter(party == "svp")

other_pred_2023 <- fundamentals_2023 %>%
  filter(party == "other")

```

###RUN DYNAMIC MODEL

```{r}
N_p <- length(party_levels)
N_comp <- length(companies)

N_polls_2019 <- nrow(y_2019)
N_polls_2023 <- nrow(y_2023)

N_periods_2019 <- length(all_weeks_2019)
N_periods_2023 <- length(all_weeks_2023)

comp_id_2019 <- poll_metadata_2019$company_id
comp_id_2023 <- poll_metadata_2023$company_id

date_2019 <- poll_metadata_2019$time_id
date_2023 <- poll_metadata_2023$time_id

#y_2019 defined above
#y_2023 defined above

#vE defined above
#vE_sd defined above

```

```{r}
#list of data for stan
dynamic_stan_2019 <- list(
  N_p = N_p,
  N_comp = N_comp,
  N_polls = N_polls_2019,
  N_periods = N_periods_2019,
  comp_id = comp_id_2019,
  date = date_2019,
  y = y_2019,
  vE = vE_2019,
  vE_sd = vE_sd_2019
)

dynamic_stan_2023 <- list(
  N_p = N_p,
  N_comp = N_comp,
  N_polls = N_polls_2023,
  N_periods = N_periods_2023,
  comp_id = comp_id_2023,
  date = date_2023,
  y = y_2023,
  vE = vE_2023,
  vE_sd = vE_sd_2023
)
```

```{r
}
dynamic_fit_2019_1m <- stan(
  file = "dynamic_model.stan",
  data = dynamic_stan_2019,
  iter = 4000,
  warmup = 2000,
  chains = 4,
  seed = 266,
  refresh = 500
)

saveRDS(dynamic_fit_2019_1m, "stan_dynamic_fit_2019_1m.rds")

```

```{r
}
dynamic_fit_2023_1m <- stan(
  file = "dynamic_model.stan",
  data = dynamic_stan_2023,
  iter = 2000,
  chains = 4,
  seed = 266,
  refresh = 500
)

saveRDS(dynamic_fit_2023_1m, "stan_dynamic_fit_2023_1m.rds")
```


###Extract Data

```{r}
#read stan data without running code again
#
dynamic_fit_2019_1m <- readRDS("stan_dynamic_fit_2019_1m.rds")
#
dynamic_fit_2023_1m <- readRDS("stan_dynamic_fit_2023_1m.rds")
```


####2019
```{r}
dynamic_forecast_2019_1m <- rstan::extract(dynamic_fit_2019_1m, "forecast_final")$forecast_final
dynamic_trajectory_2019_1m <- rstan::extract(dynamic_fit_2019_1m, "trajectory")$trajectory

N_iter_2019_1m <- dim(dynamic_trajectory_2019_1m)[1]
N_time_2019_1m <- dim(dynamic_trajectory_2019_1m)[2]
N_parties_2019_1m <- dim(dynamic_trajectory_2019_1m)[3]

```

```{r}
df_dynamic_forecast_2019_1m <- as.data.frame(dynamic_forecast_2019_1m)
colnames(df_dynamic_forecast_2019_1m) = party_levels

```

```{r}
df_dynamic_trajectory_2019_1m <- as.data.frame.table(dynamic_trajectory_2019_1m, responseName = "share") 
df_dynamic_trajectory_2019_1m <- df_dynamic_trajectory_2019_1m %>%
  mutate(
    iter = as.numeric(iterations),
    time_id = as.numeric(Var2),
    party_id = as.numeric(Var3)
  ) %>%
  mutate(
    party = party_levels[party_id],
    party = factor(party, levels = party_levels),
    date = all_weeks_2019[time_id]
  ) %>%
  select(iter, time_id, date, party, share)

```

```{r}
df_dynamic_summary_2019_1m <- df_dynamic_trajectory_2019_1m %>%
  group_by(date, party) %>%
  summarise(
    mean_pred = mean(share),
    sd_pred = sd(share),
    lower_ci = quantile(share, 0.025),
    upper_ci = quantile(share, 0.975),
    .groups = "drop"
  )

write_csv(df_dynamic_summary_2019_1m, "dynamic_predictions_2019_1m.csv")

```

```{r}
#house effects
he_dyn_2019_1m <- rstan::extract(dynamic_fit_2019_1m, "house_effect_raw")$house_effect_raw
df_he_2019_1m <- as.data.frame.table(he_dyn_2019_1m, responseName = "bias") %>%
  mutate(
    iteration = as.numeric(iterations),
    company_id = as.numeric(Var2),
    party_id = as.numeric(Var3),
    company = c("tamedia", "fors")[company_id],
    party = c("fdp", "cvp", "sp", "svp")[party_id]
  ) %>%
  select(iteration, company, party, bias)

df_he_summary_2019_1m <- df_he_2019_1m %>%
  group_by(company, party) %>%
  summarise(
    mean_bias = mean(bias),
    lower_ci = quantile(bias, 0.025),
    upper_ci = quantile(bias, 0.975),
    .groups = "drop"
  ) %>%
  mutate(
    mean_bias = mean_bias *100,
    lower_ci = lower_ci *100,
    upper_ci = upper_ci *100
  )

print(df_he_summary_2019_1m)

write_csv(df_he_summary_2019_1m, "dynamic_house_effects_2019_1m.csv")

```
These biases are relative to the sotomo polls.


```{r}
#print predictions dynamic model vs predictions fundamentals model vs polls vs election outcomes
df_dyn_result_2019_1m <- df_dynamic_summary_2019_1m[296:300, ]
df_dyn_result_2019_1m$party <- factor(df_dyn_result_2019_1m$party, levels = party_levels)

cat("---Dynamic Model Prediction 2019---\n")
print(df_dyn_result_2019_1m)

cat("\n---Fundamentals Model Prediction 2019---\n")
print(fundamentals_2019)

cat("\n---Poll Prediction 2019---\n")
print(tail(ewma_2019_1m, 1))

cat("\n---Election Results 2019---\n")
print(results_2019)

```


####2023
```{r}
dynamic_forecast_2023_1m <- rstan::extract(dynamic_fit_2023_1m, "forecast_final")$forecast_final
dynamic_trajectory_2023_1m <- rstan::extract(dynamic_fit_2023_1m, "trajectory")$trajectory

N_iter_2023_1m <- dim(dynamic_trajectory_2023_1m)[1]
N_time_2023_1m <- dim(dynamic_trajectory_2023_1m)[2]
N_parties_2023_1m <- dim(dynamic_trajectory_2023_1m)[3]

```

```{r}
df_dynamic_2023_1m <- as.data.frame(dynamic_forecast_2023_1m)
colnames(df_dynamic_2023_1m) = party_levels

```

```{r}
df_dynamic_trajectory_2023_1m <- as.data.frame.table(dynamic_trajectory_2023_1m, responseName = "share")
df_dynamic_trajectory_2023_1m <- df_dynamic_trajectory_2023_1m %>%
  mutate(
    iter = as.numeric(iterations),
    time_id = as.numeric(Var2),
    party_id = as.numeric(Var3)
  ) %>%
  mutate(
    party = party_levels[party_id],
    party = factor(party, levels = party_levels),
    date = all_weeks_2023[time_id]
  ) %>%
  select(iter, time_id, date, party, share)
```

```{r}
df_dynamic_summary_2023_1m <- df_dynamic_trajectory_2023_1m %>%
  group_by(date, party) %>%
  summarise(
    mean_pred = mean(share),
    sd_pred = sd(share),
    lower_ci = quantile(share, 0.025),
    upper_ci = quantile(share, 0.975),
    .groups = "drop"
  )

write_csv(df_dynamic_summary_2023_1m, "dynamic_predictions_2023_1m.csv")

```

```{r}
#house effects
he_dyn_2023_1m <- rstan::extract(dynamic_fit_2023_1m, "house_effect_raw")$house_effect_raw
df_he_2023_1m <- as.data.frame.table(he_dyn_2023_1m, responseName = "bias") %>%
  mutate(
    iteration = as.numeric(iterations),
    company_id = as.numeric(Var2),
    party_id = as.numeric(Var3),
    company = c("tamedia", "sotomo")[company_id],
    party = c("fdp", "cvp", "sp", "svp")[party_id]
  ) %>%
  select(iteration, company, party, bias)

df_he_summary_2023_1m <- df_he_2023_1m %>%
  group_by(company, party) %>%
  summarise(
    mean_bias = mean(bias),
    lower_ci = quantile(bias, 0.025),
    upper_ci = quantile(bias, 0.975),
    .groups = "drop"
  ) %>%
  mutate(
    mean_bias = mean_bias *100,
    lower_ci = lower_ci *100,
    upper_ci = upper_ci *100
  )

print(df_he_summary_2023_1m)

write_csv(df_he_summary_2023_1m, "dynamic_house_effects_2023_1m.csv")

```
Recall biases are relative to sotomo polls.


```{r}
#print predictions dynamic model vs predictions fundamentals model vs polls vs election outcomes

cat("---Dynamic Model Prediction 2023---\n")
df_dyn_result_2023_1m <- df_dynamic_summary_2023_1m[296:300, ]
print(df_dyn_result_2023_1m)

cat("\n---Fundamentals Model Prediction 2023---\n")
print(fundamentals_2023)

cat("\n---Poll Prediction 2023---\n")
print(tail(ewma_2023_1m, 1))

cat("\n---Election Results 2023---\n")
print(results_2023)
```


###Plots

####2019


```{r}
p_ci_2019_1m <- ggplot(df_dyn_result_2019_1m, aes(x = party)) +
  #actual result
  geom_rect(data = results_2019,
            aes(width = 0.4,
                ymin = 0, ymax = vote_share / 100, fill = party, alpha = 0.6),
            color = NA, show.legend = FALSE) +
  #95% CI
  geom_rect(aes(width = 0.3,
                ymin = lower_ci, ymax = upper_ci),
            fill = "gray50", alpha = 0.6, color = NA) +
  #mean
  geom_point(aes(y = mean_pred),
             shape = 21, fill = "white", color = "black", size = 3, stroke = 0.5) +
  #scale y to percentage
  scale_y_continuous(labels = scales::percent_format(accuracy = 1),
                     limits = c(0, 0.4),
                     expand = expansion(mult = c(0, 0.1))) +
  #party colors
  scale_fill_manual(values = party_colors) +
  #values next to bars
  geom_text(aes(y = mean_pred, label = sprintf("%.1f", mean_pred * 100)),
            hjust = -0.9, size = 4.5, color = "gray40") +
  geom_text(data = results_2019, aes(y = vote_share / 100, label = sprintf("%.1f", vote_share)),
            hjust = 1.9, size = 4.5, color = "black") +
  #party names lowercase to capital
  scale_x_discrete(labels = party_labels) +
  labs(
    x = "Party",
    y = "Vote Share"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_line(color = "gray90", linetype = "dashed"),
    panel.grid.major.x = element_blank(),
    axis.text.x = element_text(face = "bold", size = 12),
    axis.text.y = element_text(size = 12),
    axis.ticks.x = element_blank(),
    plot.margin = margin(t = 10, r = 10, b = 10, l = 10, unit = "pt")
  )

print(p_ci_2019_1m)

ggsave("Dynamic Prediction 1 Month Prior (2019).png", p_ci_2019_1m, width = 8, height = 5, dpi = 200, units = "in")

```



Check if the results fall within the CI
```{r}
check_dyn_2019_1m <- left_join(df_dyn_result_2019_1m, results_2019, by = "party") %>%
  mutate(
    within_ci = vote_share/100 >= lower_ci & vote_share/100 <= upper_ci
  )

cat("\nCheck whether the results lie within the 95% CI:\n")
print(check_dyn_2019_1m %>% select(party, vote_share, lower_ci, upper_ci, within_ci))

```

```{r}
p_all_dyn_2019_1m <- ggplot(df_dynamic_summary_2019_1m %>% filter(date >= start_2019), aes(x=date, y=mean_pred*100, color=party))+
  geom_line(linewidth=1)+
  geom_ribbon(aes(ymin=lower_ci*100, ymax=upper_ci*100, fill=party), alpha=0.2, color=NA)+
  #geom_point(data=SPD_2019 %>% filter(polling_period >= start_2019),
             #aes(x=polling_period, y=poll_vote_share, group="polling company"),
             #alpha=0.3, size=0.5, color="gray40")+
  scale_color_manual(name = "", values=c("dodgerblue4", "darkorange3", "red3", "darkgreen", "plum3"), labels = party_labels)+
  scale_fill_manual(name = "", values=party_colors, labels = party_labels)+
  scale_y_continuous(limits = c(5, 40), breaks = seq(10, 40, by = 10), minor_breaks = seq(5, 40, by = 5), expand = expansion(0))+
  labs(
    x="Date",
    y="Vote Share (%)"
  )+
  theme_minimal()+
  theme(
    legend.position = "bottom",
    plot.margin = margin(t = 10, r = 10, b = 10, l = 10, unit = "pt"))



print(p_all_dyn_2019_1m)

ggsave("Dynamic Forecast Trajectory 1 Month Prior (2019).png", p_all_dyn_2019_1m, width = 8, height = 4, dpi = 200, units = "in")

```




#####BY PARTY
Data
```{r}
#fdp 2019

fdp_result_2019 <- results_2019$vote_share[1]

fdp_polls_2019_1m <- SPD_2019 %>%
  filter(party == "fdp") %>%
  arrange(polling_period) %>%
  filter(polling_period <= cutoff_1m_2019)

fdp_ewma_2019_1m <- ewma_2019_1m %>%
  select(polling_period, ewma_fdp) %>%
  rename(ewma = ewma_fdp)

#fdp_pred_2019 = fundamentals prediction

fdp_dyn_2019_1m <- df_dynamic_summary_2019_1m %>%
  filter(party == "fdp") %>%
  mutate(
    mean_pred = mean_pred * 100,
    sd_pred = sd_pred * 100,
    lower_ci = lower_ci * 100,
    upper_ci = upper_ci * 100
  )




#cvp 2019

cvp_result_2019 <- results_2019$vote_share[2]

cvp_polls_2019_1m <- SPD_2019 %>%
  filter(party == "cvp") %>%
  arrange(polling_period) %>%
  filter(polling_period <= cutoff_1m_2019)

cvp_ewma_2019_1m <- ewma_2019_1m %>%
  select(polling_period, ewma_cvp) %>%
  rename(ewma = ewma_cvp)

#cvp_pred_2019 = fundamentals prediction

cvp_dyn_2019_1m <- df_dynamic_summary_2019_1m %>%
  filter(party == "cvp") %>%
  mutate(
    mean_pred = mean_pred * 100,
    sd_pred = sd_pred * 100,
    lower_ci = lower_ci * 100,
    upper_ci = upper_ci * 100
  )




#sp 2019

sp_result_2019 <- results_2019$vote_share[3]

sp_polls_2019_1m <- SPD_2019 %>%
  filter(party == "sp") %>%
  arrange(polling_period) %>%
  filter(polling_period <= cutoff_1m_2019)

sp_ewma_2019_1m <- ewma_2019_1m %>%
  select(polling_period, ewma_sp) %>%
  rename(ewma = ewma_sp)

#sp_pred_2019 = fundamentals prediction

sp_dyn_2019_1m <- df_dynamic_summary_2019_1m %>%
  filter(party == "sp") %>%
  mutate(
    mean_pred = mean_pred * 100,
    sd_pred = sd_pred * 100,
    lower_ci = lower_ci * 100,
    upper_ci = upper_ci * 100
  )



#svp 2019

svp_result_2019 <- results_2019$vote_share[4]

svp_polls_2019_1m <- SPD_2019 %>%
  filter(party == "svp") %>%
  arrange(polling_period) %>%
  filter(polling_period <= cutoff_1m_2019)

svp_ewma_2019_1m <- ewma_2019_1m %>%
  select(polling_period, ewma_svp) %>%
  rename(ewma = ewma_svp)

#svp_pred_2019 = fundamentals prediction

svp_dyn_2019_1m <- df_dynamic_summary_2019_1m %>%
  filter(party == "svp") %>%
  mutate(
    mean_pred = mean_pred * 100,
    sd_pred = sd_pred * 100,
    lower_ci = lower_ci * 100,
    upper_ci = upper_ci * 100
  )



#other 2019

other_result_2019 <- results_2019$vote_share[5]

other_polls_2019_1m <- SPD_2019 %>%
  filter(party == "other") %>%
  arrange(polling_period) %>%
  filter(polling_period <= cutoff_1m_2019)

other_ewma_2019_1m <- ewma_2019_1m %>%
  select(polling_period, ewma_other) %>%
  rename(ewma = ewma_other)

#other_pred_2019 = fundamentals prediction

other_dyn_2019_1m <- df_dynamic_summary_2019_1m %>%
  filter(party == "other") %>%
  mutate(
    mean_pred = mean_pred * 100,
    sd_pred = sd_pred * 100,
    lower_ci = lower_ci * 100,
    upper_ci = upper_ci * 100
  )

```

Plots
```{r}
#fdp 2019
p_dyn_fdp_2019_1m <- ggplot() +
  #dynamic model ribbon (95% ci)
  geom_ribbon(
    data = fdp_dyn_2019_1m,
    aes(x = date, ymin = lower_ci, ymax = upper_ci),
    fill = party_colors["fdp"],
    alpha = 0.2,
    inherit.aes = FALSE
  ) +
  #fundamentals (95% ci)
  geom_rect(
    data = fdp_pred_2019,
    aes(xmin = start_2019-14, xmax = fdp_dyn_2019_1m$date[1], ymin = lower_ci * 100, ymax = upper_ci * 100),
    fill = "gray80",
    alpha = 0.4,
    inherit.aes = FALSE
  ) +
  #actual result
  geom_hline(
    yintercept = fdp_result_2019,
    color = party_colors["fdp"],
    linewidth = 1,
    linetype = "solid"
  ) +
  #poll points
  geom_point(
    data = fdp_polls_2019_1m %>% filter(polling_period >= start_2019),
    aes(x = polling_period, y = poll_vote_share, shape = factor(polling_company)),
    size = 2,
    color = "black"
  ) +
  #EWMA line
  geom_line(
    data = fdp_ewma_2019_1m,
    aes(x = polling_period, y = ewma),
    color = "black",
    linewidth = 0.5
  ) +
  #fundamentals mean
  geom_hline(
    yintercept = fdp_pred_2019$mean_pred * 100,
    color = "black",
    linewidth = 0.4,
    linetype = "dashed"
  ) +
  #dynamic model mean trajectory
  geom_line(
    data = fdp_dyn_2019_1m,
    aes(x = date, y = mean_pred),
    color = "dodgerblue4",
    linewidth = 0.8
  ) +
  scale_y_continuous(limits = c(5, 30), breaks = seq(5, 30, by = 5), minor_breaks = seq(5, 30, by = 2.5), expand = expansion(0)) +
  labs(
    x = "Date",
    y = "Vote Share (%)",
    shape = "Polling Company"
  ) +
  scale_shape_discrete(name = "", labels = company_labels) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    plot.title = element_text(face = "bold"),
    plot.margin = margin(t = 20, r = 20, b = 0, l = 10, unit = "pt")
  )

print(p_dyn_fdp_2019_1m)

ggsave("FDP Pre-Election Polling vs. Dynamic Forecast 1 Month Prior (2019).png", p_dyn_fdp_2019_1m, width = 6, height = 4, dpi = 200, units = "in")



#cvp 2019

p_dyn_cvp_2019_1m <- ggplot() +
  #dynamic model ribbon (95% ci)
  geom_ribbon(
    data = cvp_dyn_2019_1m,
    aes(x = date, ymin = lower_ci, ymax = upper_ci),
    fill = party_colors["cvp"],
    alpha = 0.2,
    inherit.aes = FALSE
  ) +
  #fundamentals (95% ci)
  geom_rect(
    data = cvp_pred_2019,
    aes(xmin = start_2019-14, xmax = cvp_dyn_2019_1m$date[1], ymin = lower_ci * 100, ymax = upper_ci * 100),
    fill = "gray80",
    alpha = 0.4,
    inherit.aes = FALSE
  ) +
  #actual result
  geom_hline(
    yintercept = cvp_result_2019,
    color = party_colors["cvp"],
    linewidth = 1,
    linetype = "solid"
  ) +
  #poll points
  geom_point(
    data = cvp_polls_2019_1m %>% filter(polling_period >= start_2019),
    aes(x = polling_period, y = poll_vote_share, shape = factor(polling_company)),
    size = 2,
    color = "black"
  ) +
  #EWMA line
  geom_line(
    data = cvp_ewma_2019_1m,
    aes(x = polling_period, y = ewma),
    color = "black",
    linewidth = 0.5
  ) +
  #fundamentals mean
  geom_hline(
    yintercept = cvp_pred_2019$mean_pred * 100,
    color = "black",
    linewidth = 0.4,
    linetype = "dashed"
  ) +
  #dynamic model mean trajectory
  geom_line(
    data = cvp_dyn_2019_1m,
    aes(x = date, y = mean_pred),
    color = "darkorange3",
    linewidth = 0.8
  ) +
  scale_y_continuous(limits = c(0, 25), breaks = seq(0, 25, by = 5), minor_breaks = seq(0, 25, by = 2.5), expand = expansion(0)) +
  labs(
    x = "Date",
    y = "Vote Share (%)",
    shape = "Polling Company"
  ) +
  scale_shape_discrete(name = "", labels = company_labels) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    plot.title = element_text(face = "bold"),
    plot.margin = margin(t = 20, r = 20, b = 0, l = 10, unit = "pt")
  )

print(p_dyn_cvp_2019_1m)

ggsave("CVP Pre-Election Polling vs. Dynamic Forecast 1 Month Prior (2019).png", p_dyn_cvp_2019_1m, width = 6, height = 4, dpi = 200, units = "in")



#sp 2019

p_dyn_sp_2019_1m <- ggplot() +
  #dynamic model ribbon (95% ci)
  geom_ribbon(
    data = sp_dyn_2019_1m,
    aes(x = date, ymin = lower_ci, ymax = upper_ci),
    fill = party_colors["sp"],
    alpha = 0.2,
    inherit.aes = FALSE
  ) +
  #fundamentals (95% ci)
  geom_rect(
    data = sp_pred_2019,
    aes(xmin = start_2019-14, xmax = sp_dyn_2019_1m$date[1], ymin = lower_ci * 100, ymax = upper_ci * 100),
    fill = "gray80",
    alpha = 0.4,
    inherit.aes = FALSE
  ) +
  #actual result
  geom_hline(
    yintercept = sp_result_2019,
    color = party_colors["sp"],
    linewidth = 1,
    linetype = "solid"
  ) +
  #poll points
  geom_point(
    data = sp_polls_2019_1m %>% filter(polling_period >= start_2019),
    aes(x = polling_period, y = poll_vote_share, shape = factor(polling_company)),
    size = 2,
    color = "black"
  ) +
  #EWMA line
  geom_line(
    data = sp_ewma_2019_1m,
    aes(x = polling_period, y = ewma),
    color = "black",
    linewidth = 0.5
  ) +
  #fundamentals mean
  geom_hline(
    yintercept = sp_pred_2019$mean_pred * 100,
    color = "black",
    linewidth = 0.4,
    linetype = "dashed"
  ) +
  #dynamic model mean trajectory
  geom_line(
    data = sp_dyn_2019_1m,
    aes(x = date, y = mean_pred),
    color = "red3",
    linewidth = 0.8
  ) +
  scale_y_continuous(limits = c(10, 35), breaks = seq(10, 35, by = 5), minor_breaks = seq(10, 35, by = 2.5), expand = expansion(0)) +
  labs(
    x = "Date",
    y = "Vote Share (%)",
    shape = "Polling Company"
  ) +
  scale_shape_discrete(name = "", labels = company_labels) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    plot.title = element_text(face = "bold"),
    plot.margin = margin(t = 20, r = 20, b = 0, l = 10, unit = "pt")
  )

print(p_dyn_sp_2019_1m)

ggsave("SP Pre-Election Polling vs. Dynamic Forecast 1 Month Prior (2019).png", p_dyn_sp_2019_1m, width = 6, height = 4, dpi = 200, units = "in")



#svp 2019

p_dyn_svp_2019_1m <- ggplot() +
  #dynamic model ribbon (95% ci)
  geom_ribbon(
    data = svp_dyn_2019_1m,
    aes(x = date, ymin = lower_ci, ymax = upper_ci),
    fill = party_colors["svp"],
    alpha = 0.2,
    inherit.aes = FALSE
  ) +
  #fundamentals (95% ci)
  geom_rect(
    data = svp_pred_2019,
    aes(xmin = start_2019-14, xmax = svp_dyn_2019_1m$date[1], ymin = lower_ci * 100, ymax = upper_ci * 100),
    fill = "gray80",
    alpha = 0.4,
    inherit.aes = FALSE
  ) +
  #actual result
  geom_hline(
    yintercept = svp_result_2019,
    color = party_colors["svp"],
    linewidth = 1,
    linetype = "solid"
  ) +
  #poll points
  geom_point(
    data = svp_polls_2019_1m %>% filter(polling_period >= start_2019),
    aes(x = polling_period, y = poll_vote_share, shape = factor(polling_company)),
    size = 2,
    color = "black"
  ) +
  #EWMA line
  geom_line(
    data = svp_ewma_2019_1m,
    aes(x = polling_period, y = ewma),
    color = "black",
    linewidth = 0.5
  ) +
  #fundamentals mean
  geom_hline(
    yintercept = svp_pred_2019$mean_pred * 100,
    color = "black",
    linewidth = 0.4,
    linetype = "dashed"
  ) +
  #dynamic model mean trajectory
  geom_line(
    data = svp_dyn_2019_1m,
    aes(x = date, y = mean_pred),
    color = "darkgreen",
    linewidth = 0.8
  ) +
  scale_y_continuous(limits = c(15, 40), breaks = seq(15, 40, by = 5), minor_breaks = seq(15, 40, by = 2.5), expand = expansion(0)) +
  labs(
    x = "Date",
    y = "Vote Share (%)",
    shape = "Polling Company"
  ) +
  scale_shape_discrete(name = "", labels = company_labels) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    plot.title = element_text(face = "bold"),
    plot.margin = margin(t = 20, r = 20, b = 0, l = 10, unit = "pt")
  )

print(p_dyn_svp_2019_1m)

ggsave("SVP Pre-Election Polling vs. Dynamic Forecast 1 Month Prior (2019).png", p_dyn_svp_2019_1m, width = 6, height = 4, dpi = 200, units = "in")



#other 2019

p_dyn_other_2019_1m <- ggplot() +
  #dynamic model ribbon (95% ci)
  geom_ribbon(
    data = other_dyn_2019_1m,
    aes(x = date, ymin = lower_ci, ymax = upper_ci),
    fill = party_colors["other"],
    alpha = 0.2,
    inherit.aes = FALSE
  ) +
  #fundamentals (95% ci)
  geom_rect(
    data = other_pred_2019,
    aes(xmin = start_2019-14, xmax = other_dyn_2019_1m$date[1], ymin = lower_ci * 100, ymax = upper_ci * 100),
    fill = "gray80",
    alpha = 0.4,
    inherit.aes = FALSE
  ) +
  #actual result
  geom_hline(
    yintercept = other_result_2019,
    color = party_colors["other"],
    linewidth = 1,
    linetype = "solid"
  ) +
  #poll points
  geom_point(
    data = other_polls_2019_1m %>% filter(polling_period >= start_2019),
    aes(x = polling_period, y = poll_vote_share, shape = factor(polling_company)),
    size = 2,
    color = "black"
  ) +
  #EWMA line
  geom_line(
    data = other_ewma_2019_1m,
    aes(x = polling_period, y = ewma),
    color = "black",
    linewidth = 0.5
  ) +
  #fundamentals mean
  geom_hline(
    yintercept = other_pred_2019$mean_pred * 100,
    color = "black",
    linewidth = 0.4,
    linetype = "dashed"
  ) +
  #dynamic model mean trajectory
  geom_line(
    data = other_dyn_2019_1m,
    aes(x = date, y = mean_pred),
    color = "plum3",
    linewidth = 0.8
  ) +
  scale_y_continuous(limits = c(15, 40), breaks = seq(5, 40, by = 5), minor_breaks = seq(5, 40, by = 2.5), expand = expansion(0)) +
  labs(
    x = "Date",
    y = "Vote Share (%)",
    shape = "Polling Company"
  ) +
  scale_shape_discrete(name = "", labels = company_labels) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    plot.title = element_text(face = "bold"),
    plot.margin = margin(t = 20, r = 20, b = 0, l = 10, unit = "pt")
  )

print(p_dyn_other_2019_1m)

ggsave("Other Pre-Election Polling vs. Dynamic Forecast 1 Month Prior (2019).png", p_dyn_other_2019_1m, width = 6, height = 4, dpi = 200, units = "in")

```


####2023


```{r}
p_ci_2023_1m <- ggplot(df_dyn_result_2023_1m, aes(x = party)) +
  #actual result
  geom_rect(data = results_2023,
            aes(width = 0.4,
                ymin = 0, ymax = vote_share / 100, fill = party, alpha = 0.6),
            color = NA, show.legend = FALSE) +
  #95% CI
  geom_rect(aes(width = 0.3,
                ymin = lower_ci, ymax = upper_ci),
            fill = "gray50", alpha = 0.6, color = NA) +
  #mean
  geom_point(aes(y = mean_pred),
             shape = 21, fill = "white", color = "black", size = 3, stroke = 0.5) +
  #scale y to percentage
  scale_y_continuous(labels = scales::percent_format(accuracy = 1),
                     limits = c(0, 0.4),
                     expand = expansion(mult = c(0, 0.1))) +
  #party colors
  scale_fill_manual(values = party_colors) +
  #values next to bars
  geom_text(aes(y = mean_pred, label = sprintf("%.1f", mean_pred * 100)),
            hjust = -0.9, size = 4.5, color = "gray40") +
  geom_text(data = results_2023, aes(y = vote_share / 100, label = sprintf("%.1f", vote_share)),
            hjust = 1.9, size = 4.5, color = "black") +
  #party names lowercase to capital
  scale_x_discrete(labels = party_labels) +
  labs(
    x = "Party",
    y = "Vote Share"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_line(color = "gray90", linetype = "dashed"),
    panel.grid.major.x = element_blank(),
    axis.text.x = element_text(face = "bold", size = 12),
    axis.text.y = element_text(size = 12),
    axis.ticks.x = element_blank(),
    plot.margin = margin(t = 10, r = 10, b = 10, l = 10, unit = "pt")
  )

print(p_ci_2023_1m)

ggsave("Dynamic Prediction 1 Month Prior (2023).png", p_ci_2023_1m, width = 8, height = 5, dpi = 200, units = "in")

```


Check if the results fall within the CI
```{r}
check_dyn_2023_1m <- left_join(df_dyn_result_2023_1m, results_2023, by = "party") %>%
  mutate(
    within_ci = vote_share/100 >= lower_ci & vote_share/100 <= upper_ci
  )

cat("\nCheck whether the results lie within the 95% CI:\n")
print(check_dyn_2023_1m %>% select(party, vote_share, lower_ci, upper_ci, within_ci))

```

```{r}
p_all_dyn_2023_1m <- ggplot(df_dynamic_summary_2023_1m %>% filter(date >= start_2023), aes(x=date, y=mean_pred*100, color=party))+
  geom_line(linewidth=1)+
  geom_ribbon(aes(ymin=lower_ci*100, ymax=upper_ci*100, fill=party), alpha=0.2, color=NA)+
  #geom_point(data=SPD_2023 %>% filter(polling_period >= start_2023),
             #aes(x=polling_period, y=poll_vote_share, group="polling company"),
             #alpha=0.3, size=0.5, color="gray40")+
  scale_color_manual(name = "", values=c("dodgerblue4", "darkorange3", "red3", "darkgreen", "plum3"), labels = party_labels)+
  scale_fill_manual(name = "", values=party_colors, labels = party_labels)+
  scale_y_continuous(limits = c(5, 40), breaks = seq(10, 40, by = 10), minor_breaks = seq(5, 40, by = 5), expand = expansion(0))+
  labs(
    x="Date",
    y="Vote Share (%)"
  )+
  theme_minimal()+
  theme(legend.position = "bottom",
        plot.margin = margin(t = 10, r = 10, b = 10, l = 10, unit = "pt"))



print(p_all_dyn_2023_1m)

ggsave("Dynamic Forecast Trajectory 1 Month Prior (2023).png", p_all_dyn_2023_1m, width = 8, height = 4, dpi = 200, units = "in")

```


#####BY PARTY
Data
```{r}
#fdp 2023

fdp_result_2023 <- results_2023$vote_share[1]

fdp_polls_2023_1m <- SPD_2023 %>%
  filter(party == "fdp") %>%
  arrange(polling_period)%>%
  filter(polling_period <= cutoff_1m_2023)

fdp_ewma_2023_1m <- ewma_2023_1m %>%
  select(polling_period, ewma_fdp) %>%
  rename(ewma = ewma_fdp)

#fdp_pred_2023 = fundamentals prediction

fdp_dyn_2023_1m <- df_dynamic_summary_2023_1m %>%
  filter(party == "fdp") %>%
  mutate(
    mean_pred = mean_pred * 100,
    sd_pred = sd_pred * 100,
    lower_ci = lower_ci * 100,
    upper_ci = upper_ci *100
  )



#cvp 2023

cvp_result_2023 <- results_2023$vote_share[2]

cvp_polls_2023_1m <- SPD_2023 %>%
  filter(party == "cvp") %>%
  arrange(polling_period) %>%
  filter(polling_period <= cutoff_1m_2023)

cvp_ewma_2023_1m <- ewma_2023_1m %>%
  select(polling_period, ewma_cvp) %>%
  rename(ewma = ewma_cvp)

#cvp_pred_2023 = fundamentals prediction

cvp_dyn_2023_1m <- df_dynamic_summary_2023_1m %>%
  filter(party == "cvp") %>%
  mutate(
    mean_pred = mean_pred * 100,
    sd_pred = sd_pred * 100,
    lower_ci = lower_ci * 100,
    upper_ci = upper_ci *100
  )



#sp 2023

sp_result_2023 <- results_2023$vote_share[3]

sp_polls_2023_1m <- SPD_2023 %>%
  filter(party == "sp") %>%
  arrange(polling_period) %>%
  filter(polling_period <= cutoff_1m_2023)

sp_ewma_2023_1m <- ewma_2023_1m %>%
  select(polling_period, ewma_sp) %>%
  rename(ewma = ewma_sp)

#sp_pred_2023 = fundamentals prediction

sp_dyn_2023_1m <- df_dynamic_summary_2023_1m %>%
  filter(party == "sp") %>%
  mutate(
    mean_pred = mean_pred * 100,
    sd_pred = sd_pred * 100,
    lower_ci = lower_ci * 100,
    upper_ci = upper_ci *100
  )



#svp 2023

svp_result_2023 <- results_2023$vote_share[4]

svp_polls_2023_1m <- SPD_2023 %>%
  filter(party == "svp") %>%
  arrange(polling_period) %>%
  filter(polling_period <= cutoff_1m_2023)

svp_ewma_2023_1m <- ewma_2023_1m %>%
  select(polling_period, ewma_svp) %>%
  rename(ewma = ewma_svp)

#svp_pred_2023 = fundamentals prediction

svp_dyn_2023_1m <- df_dynamic_summary_2023_1m %>%
  filter(party == "svp") %>%
  mutate(
    mean_pred = mean_pred * 100,
    sd_pred = sd_pred * 100,
    lower_ci = lower_ci * 100,
    upper_ci = upper_ci *100
  )



#other 2023

other_result_2023 <- results_2023$vote_share[5]

other_polls_2023_1m <- SPD_2023 %>%
  filter(party == "other") %>%
  arrange(polling_period) %>%
  filter(polling_period <= cutoff_1m_2023)

other_ewma_2023_1m <- ewma_2023_1m %>%
  select(polling_period, ewma_other) %>%
  rename(ewma = ewma_other)

#other_pred_2023 = fundamentals prediction

other_dyn_2023_1m <- df_dynamic_summary_2023_1m %>%
  filter(party == "other") %>%
  mutate(
    mean_pred = mean_pred * 100,
    sd_pred = sd_pred * 100,
    lower_ci = lower_ci * 100,
    upper_ci = upper_ci *100
  )


```


Plots
```{r}
#fdp 2023

p_dyn_fdp_2023_1m <- ggplot() +
  #dynamic model ribbon (95% ci)
  geom_ribbon(
    data = fdp_dyn_2023_1m,
    aes(x = date, ymin = lower_ci, ymax = upper_ci),
    fill = party_colors["fdp"],
    alpha = 0.2,
    inherit.aes = FALSE
  ) +
  #fundamentals (95% ci)
  geom_rect(
    data = fdp_pred_2023,
    aes(xmin = start_2023 -14, xmax = fdp_dyn_2023_1m$date[1], ymin = lower_ci * 100, ymax = upper_ci * 100),
    fill = "gray80",
    alpha = 0.4,
    inherit.aes = FALSE
  ) +
  #actual result
  geom_hline(
    yintercept = fdp_result_2023,
    color = party_colors["fdp"],
    linewidth = 1,
    linetype = "solid"
  ) +
  #poll points
  geom_point(
    data = fdp_polls_2023_1m %>% filter(polling_period >= start_2023),
    aes(x = polling_period, y = poll_vote_share, shape = factor(polling_company)),
    size = 2,
    color = "black"
  ) +
  #EWMA line
  geom_line(
    data = fdp_ewma_2023_1m,
    aes(x = polling_period, y = ewma),
    color = "black",
    linewidth = 0.5
  ) +
  #fundamentals mean
  geom_hline(
    yintercept = fdp_pred_2023$mean_pred * 100,
    color = "black",
    linewidth = 0.4,
    linetype = "dashed"
  ) +
  #dynamic model mean trajectory
  geom_line(
    data = fdp_dyn_2023_1m,
    aes(x = date, y = mean_pred),
    color = "dodgerblue4",
    linewidth = 0.8
  ) +
  scale_y_continuous(limits = c(5, 30), breaks = seq(5, 30, by = 5), minor_breaks = seq(5, 30, by = 2.5), expand = expansion(0)) +
  labs(
    x = "Date",
    y = "Vote Share (%)",
    shape = "Polling Company"
  ) +
  scale_shape_discrete(name = "", labels = company_labels) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    plot.title = element_text(face = "bold"),
    plot.margin = margin(t = 20, r = 20, b = 0, l = 10, unit = "pt")
  )

print(p_dyn_fdp_2023_1m)

ggsave("FDP Pre-Election Polling vs. Dynamic Forecast 1 Month Prior (2023).png", p_dyn_fdp_2023_1m, width = 6, height = 4, dpi = 200, units = "in")


#cvp 2023

p_dyn_cvp_2023_1m <- ggplot() +
  #dynamic model ribbon (95% ci)
  geom_ribbon(
    data = cvp_dyn_2023_1m,
    aes(x = date, ymin = lower_ci, ymax = upper_ci),
    fill = party_colors["cvp"],
    alpha = 0.2,
    inherit.aes = FALSE
  ) +
  #fundamentals (95% ci)
  geom_rect(
    data = cvp_pred_2023,
    aes(xmin = start_2023 -14, xmax = cvp_dyn_2023_1m$date[1], ymin = lower_ci * 100, ymax = upper_ci * 100),
    fill = "gray80",
    alpha = 0.4,
    inherit.aes = FALSE
  ) +
  #actual result
  geom_hline(
    yintercept = cvp_result_2023,
    color = party_colors["cvp"],
    linewidth = 1,
    linetype = "solid"
  ) +
  #poll points
  geom_point(
    data = cvp_polls_2023_1m %>% filter(polling_period >= start_2023),
    aes(x = polling_period, y = poll_vote_share, shape = factor(polling_company)),
    size = 2,
    color = "black"
  ) +
  #EWMA line
  geom_line(
    data = cvp_ewma_2023_1m,
    aes(x = polling_period, y = ewma),
    color = "black",
    linewidth = 0.5
  ) +
  #fundamentals mean
  geom_hline(
    yintercept = cvp_pred_2023$mean_pred * 100,
    color = "black",
    linewidth = 0.4,
    linetype = "dashed"
  ) +
  #dynamic model mean trajectory
  geom_line(
    data = cvp_dyn_2023_1m,
    aes(x = date, y = mean_pred),
    color = "darkorange3",
    linewidth = 0.8
  ) +
  scale_y_continuous(limits = c(0, 25), breaks = seq(0, 25, by = 5), minor_breaks = seq(0, 25, by = 2.5), expand = expansion(0)) +
  labs(
    x = "Date",
    y = "Vote Share (%)",
    shape = "Polling Company"
  ) +
  scale_shape_discrete(name = "", labels = company_labels) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    plot.title = element_text(face = "bold"),
    plot.margin = margin(t = 20, r = 20, b = 0, l = 10, unit = "pt")
  )

print(p_dyn_cvp_2023_1m)

ggsave("CVP Pre-Election Polling vs. Dynamic Forecast 1 Month Prior (2023).png", p_dyn_cvp_2023_1m, width = 6, height = 4, dpi = 200, units = "in")




#sp 2023

p_dyn_sp_2023_1m <- ggplot() +
  #dynamic model ribbon (95% ci)
  geom_ribbon(
    data = sp_dyn_2023_1m,
    aes(x = date, ymin = lower_ci, ymax = upper_ci),
    fill = party_colors["sp"],
    alpha = 0.2,
    inherit.aes = FALSE
  ) +
  #fundamentals (95% ci)
  geom_rect(
    data = sp_pred_2023,
    aes(xmin = start_2023 -14, xmax = sp_dyn_2023_1m$date[1], ymin = lower_ci * 100, ymax = upper_ci * 100),
    fill = "gray80",
    alpha = 0.4,
    inherit.aes = FALSE
  ) +
  #actual result
  geom_hline(
    yintercept = sp_result_2023,
    color = party_colors["sp"],
    linewidth = 1,
    linetype = "solid"
  ) +
  #poll points
  geom_point(
    data = sp_polls_2023_1m %>% filter(polling_period >= start_2023),
    aes(x = polling_period, y = poll_vote_share, shape = factor(polling_company)),
    size = 2,
    color = "black"
  ) +
  #fundamentals mean
  geom_hline(
    yintercept = sp_pred_2023$mean_pred * 100,
    color = "black",
    linewidth = 0.4,
    linetype = "dashed"
  ) +
  #EWMA line
  geom_line(
    data = sp_ewma_2023_1m,
    aes(x = polling_period, y = ewma),
    color = "black",
    linewidth = 0.5
  ) +
  #dynamic model mean trajectory
  geom_line(
    data = sp_dyn_2023_1m,
    aes(x = date, y = mean_pred),
    color = "red3",
    linewidth = 0.8
  ) +
  scale_y_continuous(limits = c(10, 35), breaks = seq(10, 35, by = 5), minor_breaks = seq(10, 35, by = 2.5), expand = expansion(0)) +
  labs(
    x = "Date",
    y = "Vote Share (%)",
    shape = "Polling Company"
  ) +
  scale_shape_discrete(name = "", labels = company_labels) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    plot.title = element_text(face = "bold"),
    plot.margin = margin(t = 20, r = 20, b = 0, l = 10, unit = "pt")
  )

print(p_dyn_sp_2023_1m)

ggsave("SP Pre-Election Polling vs. Dynamic Forecast 1 Month Prior (2023).png", p_dyn_sp_2023_1m, width = 6, height = 4, dpi = 200, units = "in")




#svp 2023

p_dyn_svp_2023_1m <- ggplot() +
  #dynamic model ribbon (95% ci)
  geom_ribbon(
    data = svp_dyn_2023_1m,
    aes(x = date, ymin = lower_ci, ymax = upper_ci),
    fill = party_colors["svp"],
    alpha = 0.2,
    inherit.aes = FALSE
  ) +
  #fundamentals (95% ci)
  geom_rect(
    data = svp_pred_2023,
    aes(xmin = start_2023 -14, xmax = svp_dyn_2023_1m$date[1], ymin = lower_ci * 100, ymax = upper_ci * 100),
    fill = "gray80",
    alpha = 0.4,
    inherit.aes = FALSE
  ) +
  #poll points
  geom_point(
    data = svp_polls_2023_1m %>% filter(polling_period >= start_2023),
    aes(x = polling_period, y = poll_vote_share, shape = factor(polling_company)),
    size = 2,
    color = "black"
  ) +
  #actual result
  geom_hline(
    yintercept = svp_result_2023,
    color = party_colors["svp"],
    linewidth = 1,
    linetype = "solid"
  ) +
  #EWMA line
  geom_line(
    data = svp_ewma_2023_1m,
    aes(x = polling_period, y = ewma),
    color = "black",
    linewidth = 0.5
  ) +
  #fundamentals mean
  geom_hline(
    yintercept = svp_pred_2023$mean_pred * 100,
    color = "black",
    linewidth = 0.4,
    linetype = "dashed"
  ) +
  #dynamic model mean trajectory
  geom_line(
    data = svp_dyn_2023_1m,
    aes(x = date, y = mean_pred),
    color = "darkgreen",
    linewidth = 0.8
  ) +
  scale_y_continuous(limits = c(15, 40), breaks = seq(15, 40, by = 5), minor_breaks = seq(15, 40, by = 2.5), expand = expansion(0)) +
  labs(
    x = "Date",
    y = "Vote Share (%)",
    shape = "Polling Company"
  ) +
  scale_shape_discrete(name = "", labels = company_labels) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    plot.title = element_text(face = "bold"),
    plot.margin = margin(t = 20, r = 20, b = 0, l = 10, unit = "pt")
  )

print(p_dyn_svp_2023_1m)

ggsave("SVP Pre-Election Polling vs. Dynamic Forecast 1 Month Prior (2023).png", p_dyn_svp_2023_1m, width = 6, height = 4, dpi = 200, units = "in")



#other 2023

p_dyn_other_2023_1m <- ggplot() +
  #dynamic model ribbon (95% ci)
  geom_ribbon(
    data = other_dyn_2023_1m,
    aes(x = date, ymin = lower_ci, ymax = upper_ci),
    fill = party_colors["other"],
    alpha = 0.2,
    inherit.aes = FALSE
  ) +
  #fundamentals (95% ci)
  geom_rect(
    data = other_pred_2023,
    aes(xmin = start_2023 -14, xmax = other_dyn_2023_1m$date[1], ymin = lower_ci * 100, ymax = upper_ci * 100),
    fill = "gray80",
    alpha = 0.4,
    inherit.aes = FALSE
  ) +
  #actual result
  geom_hline(
    yintercept = other_result_2023,
    color = party_colors["other"],
    linewidth = 1,
    linetype = "solid"
  ) +
  #poll points
  geom_point(
    data = other_polls_2023_1m %>% filter(polling_period >= start_2023),
    aes(x = polling_period, y = poll_vote_share, shape = factor(polling_company)),
    size = 2,
    color = "black"
  ) +
  #EWMA line
  geom_line(
    data = other_ewma_2023_1m,
    aes(x = polling_period, y = ewma),
    color = "black",
    linewidth = 0.5
  ) +
  #fundamentals mean
  geom_hline(
    yintercept = other_pred_2023$mean_pred * 100,
    color = "black",
    linewidth = 0.4,
    linetype = "dashed"
  ) +
  #dynamic model mean trajectory
  geom_line(
    data = other_dyn_2023_1m,
    aes(x = date, y = mean_pred),
    color = "plum3",
    linewidth = 0.8
  ) +
  scale_y_continuous(limits = c(10, 35), breaks = seq(10, 35, by = 5), minor_breaks = seq(10, 35, by = 2.5), expand = expansion(0)) +
  labs(
    x = "Date",
    y = "Vote Share (%)",
    shape = "Polling Company"
  ) +
  scale_shape_discrete(name = "", labels = company_labels) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    plot.title = element_text(face = "bold"),
    plot.margin = margin(t = 20, r = 20, b = 0, l = 10, unit = "pt")
  )

print(p_dyn_other_2023_1m)

ggsave("Other Pre-Election Polling vs. Dynamic Forecast 1 Month Prior (2023).png", p_dyn_other_2023_1m, width = 6, height = 4, dpi = 200, units = "in")

```