---
title: "fundamentals - predict 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Preprocessing

Load libraries and set the working directory.
```{r}
library(rstan)
library(dplyr)
library(ggplot2)
library(scales)
library(tidyr)

```

Import the election data.
```{r}
library(readr)

election_data <- read_csv("swiss election data.csv")

```

Order and colors for the parties, as well as the predictor names and the training years.
```{r}
#parties and predictors
party_levels <- c("fdp", "cvp", "sp", "svp", "other")
party_labels <- c(
  "fdp" = "FDP",
  "cvp" = "CVP",
  "sp" = "SP",
  "svp" = "SVP",
  "other" = "Other"
)

predictor_cols <- c("smi_norm", "health_insurance_norm", "immigration_norm")

party_colors <- c(
  "fdp" = "cornflowerblue",
  "cvp" = "darkorange",
  "sp" = "red2",
  "svp" = "green4",
  "other" = "plum2"
)

#predictor data is from 2003 onward
#predicting 2019 so will leave out 2023
all_years_2019 <- c(2003, 2007, 2011, 2015, 2019)
train_years_2019 <- c(2003, 2007, 2011, 2015)

```


##Data Frames

Filter the data to keep only the relevant data for predicting the 2019 election.
```{r}
df_data_2019 <- election_data %>%
  filter(year %in% all_years_2019) %>%
  mutate(
    party = factor(party, levels = party_levels),
    year = as.integer(year)
  ) %>%
  arrange(year, party)

```


##Stan Data

Number of elections that the model will consider for training are all the years from 2003 leading up to the 2019 election.
We consider 5 parties: fdp, cvp, sp, svp and "other".
The number of observations used for training is N_p * N_e.
In order to train the model using only the data from 2003 to 2015 we use N_train which is of length 4 in this case (2003, 2007, 2011, 2015).
```{r}
  N_e_2019 <-  length(all_years_2019) #5
  N_p <-  length(party_levels) #5
  N_obs_2019 <-  N_e_2019 * N_p #25
  K <-  length(predictor_cols) #3
  X_2019 <-  as.matrix(df_data_2019[, predictor_cols])
  v_obs_2019 <-  df_data_2019$vote_share
  p_id_2019 <-  as.integer(df_data_2019$party)
  e_id_2019 <-  as.integer(factor(df_data_2019$year, levels = all_years_2019))
  N_train_2019 <-  length(train_years_2019) #4
  N_pred = 1 #N_e - N_train

stan_data_2019 <- list(
  N_e = N_e_2019,
  N_p = N_p,
  N_obs = N_obs_2019,
  K = K,
  X = X_2019,
  v_obs = v_obs_2019,
  p_id = p_id_2019,
  e_id = e_id_2019,
  N_train = N_train_2019,
  N_pred = N_pred
)

```

```{r}
# Sanity Check before running Stan
cat("=== Data Dimension Check ===\n")
cat("N_e (Elections):", N_e_2019, "\n")
cat("N_p (Parties):", N_p, "\n")
cat("N_obs (Total Rows):", N_obs_2019, "\n")
cat("X Matrix Dimensions:", dim(X_2019), "\n") # Should be 25 3
cat("v_obs Length:", length(v_obs_2019), "\n") # Should be 25
cat("N_train:", N_train_2019, "\n")
cat("N_pred:", N_pred, "\n")

```


##Run Fundamentals

Run the fundamentals code with the data needed to predict the 2019 election.
```{r}
fit_2019 <- stan(
  file = "fundamentals.stan",
  data = stan_data_2019,
  iter = 2000,
  chains = 4,
  seed = 266,
  refresh = 500 #show progress every 500 iterations
)

saveRDS(fit_2019, "fundamentals_fit_2019.rds")

```


##Extract Predictions 2019

```{r}
#read stan fit without running code again
#fit_2019 <- readRDS("fundamentals_fit_2019.rds")
```


The predicted samples v_pred are extracted from Stan.
```{r}
#v_pred is [N_pred * N_p] so 5 values per iteration
sims_2019 <- rstan::extract(fit_2019, "v_pred")$v_pred

#reshape to [iterations, parties]
dim(sims_2019) <- c(dim(sims_2019)[1], N_p)

```


##Data Frames

Create a data frame for plotting.
```{r}
df_pred_2019 <-  as.data.frame(sims_2019)
colnames(df_pred_2019) = party_levels

df_pred_long_2019 <- df_pred_2019 %>%
  mutate(iteration = row_number()) %>%
  pivot_longer(cols = -iteration, names_to = "party", values_to = "pred_share") %>%
  mutate(party = factor(party, levels = party_levels))
```

Create a data frame for statistics.
```{r}
df_stats_2019 <- df_pred_long_2019 %>%
  group_by(party) %>%
  summarise(
    mean_pred = mean(pred_share),
    sd_pred = sd(pred_share),
    lower_ci = quantile(pred_share, 0.025),
    upper_ci = quantile(pred_share, 0.975),
    .groups = "drop"
  ) %>%
  mutate(year = 2019,
         party = as.character(party))

#save as csv to use in other rmds
write_csv(df_stats_2019, "fundamentals_predictions_2019.csv")

```

Create a data frame with the actual results of the 2019 election so can plot these to compare with our prediction.
```{r}
df_results_2019 <- df_data_2019 %>%
  filter(year == 2019) %>%
  select(party, vote_share)

```


##Statistics Prediction

Print out the statistics of the prediction.
```{r}
cat("--- 2019 Election PRediction Results ---\n")
print(df_stats_2019)
cat("\nActual 2019 Results:\n")
print(df_results_2019)

```

Check if the results fall within the CI
```{r}
df_check_2019 <- left_join(df_stats_2019, df_results_2019, by = "party") %>%
  mutate(
    within_ci = vote_share >= lower_ci & vote_share <= upper_ci
  )

cat("\nCheck whether the actual results lie within the 95% CI:\n")
print(df_check_2019 %>% select(party, vote_share, lower_ci, upper_ci, within_ci))

```
See df_check_2019 for more accurate values of the CI - will see that the lower bound is slightly above 0.125 and therefore the vote_share 0.125 does not lie within the interval (by not even 0.0005) and we get the output FALSE.


##Plot Prediction

Plot the 95% CI for each party and overlay the actual result.
```{r
p_2019 <- ggplot(df_stats_2019, aes(x = party, y = mean_pred)) +
  #95% CI
  geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci),
                width = 0.2,
                color = "gray40",
                linewidth = 1) +
  #mean prediction point
  geom_point(size = 4, color = "gray30", shape = 21, fill ="white") +
  #actual result
  geom_point(data = df_results_2019,
             aes(x = party, y = vote_share),
             shape = 4,
             size = 5,
             color = "red3") +
  #scale y to percentage
  scale_y_continuous(labels = scales::percent_format(accuracy = 1),
                     limits = c(0, 0.4),
                     expand = expansion(mult = c(0, 0.1))) +
  #party names lowercase to capital
  scale_x_discrete(labels = party_labels) +
  labs(
    title = "Posterior Predictive Check (2019)",
    x = "Party",
    y = "Vote Share"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(face = "bold", size = 12, color = party_colors),
    axis.text.y = element_text(size = 12)
  )

print(p_2019)

ggsave("Posterior Predictive Check (2019).png", p_2019, width = 6, height = 4, dpi = 200, units = "in")

```

```{r}
p_2019 <- ggplot(df_stats_2019, aes(x = party)) +
  #actual result
  geom_rect(data = df_results_2019,
            aes(width = 0.4,
                ymin = 0, ymax = vote_share, fill = party, alpha = 0.6),
            color = NA, show.legend = FALSE) +
  #95% CI
  geom_rect(aes(width = 0.3,
                ymin = lower_ci, ymax = upper_ci),
            fill = "gray50", alpha = 0.6, color = NA) +
  #mean
  geom_point(aes(y = mean_pred),
             shape = 21, fill = "white", color = "black", size = 3, stroke = 0.5) +
  #scale y to percentage
  scale_y_continuous(labels = scales::percent_format(accuracy = 1),
                     limits = c(0, 0.4),
                     expand = expansion(mult = c(0, 0.1))) +
  #party colors
  scale_fill_manual(values = party_colors) +
  #values next to bars
  geom_text(aes(y = mean_pred, label = sprintf("%.1f", mean_pred * 100)),
            hjust = -0.9, size = 4.5, color = "gray40") +
  geom_text(data = df_results_2019, aes(y = vote_share, label = sprintf("%.1f", vote_share * 100)),
            hjust = 1.9, size = 4.5, color = "black") +
  #party names lowercase to capital
  scale_x_discrete(labels = party_labels) +
  labs(
    x = "Party",
    y = "Vote Share"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_line(color = "gray90", linetype = "dashed"),
    panel.grid.major.x = element_blank(),
    axis.text.x = element_text(face = "bold", size = 12),
    axis.text.y = element_text(size = 12),
    axis.ticks.x = element_blank(),
    plot.margin = margin(t = 10, r = 10, b = 10, l = 10, unit = "pt")
  )

print(p_2019)

ggsave("Posterior Predictive Check (2019).png", p_2019, width = 8, height = 5, dpi = 200, units = "in")

```


##Plot Posterior Distribution of the Betas

See how each predictor affects the vote share of each party.
Positive: attracts voters, Negative: repels voters, Zero: no direct effect (vote share could increase or decrease due to direct effects on other parties)

###Data Frame Betas

Extract the posterior samples - this time all, not only the vote share predictions!
```{r}
#extract posterior samples
posterior_samples_2019 <- rstan::extract(fit_2019)

```

Process beta_0 (Intercepts): create a data frame with only the beta_0s
```{r}
#process beta_0 (intercepts)
df_beta0_2019 <- as.data.frame(posterior_samples_2019$beta_0)
colnames(df_beta0_2019) <- party_levels

df_beta0_long_2019 <- df_beta0_2019 %>%
  mutate(iteration = row_number()) %>%
  pivot_longer(cols = -iteration, names_to = "party", values_to = "value") %>%
  mutate(param_type = "Intercept (beta_0)", party = factor(party, levels = party_levels))

df_beta0_wide_2019 <- df_beta0_long_2019 %>%
  select(iteration, party, value) %>%
  pivot_wider(names_from = party, values_from = value)

df_base_shares_2019 <- df_beta0_wide_2019 %>%
  mutate(across(all_of(party_levels), exp, .names = "alpha_{.col}")) %>%
  rowwise() %>%
  mutate(
    total_alpha = sum(c_across(starts_with("alpha_"))),
    across(
      starts_with("alpha_"),
      ~ .x / total_alpha,
      .names = "share_{.col}"
    )
  ) %>%
  ungroup()

df_beta0_share_long_2019 <- df_base_shares_2019 %>%
  select(iteration, starts_with("share_alpha_")) %>%
  pivot_longer(
    cols = -iteration,
    names_to = "party",
    values_to = "base_vote_share"
  ) %>%
  mutate(
    party = gsub("share_alpha_", "", party),
    party = factor(party, levels = party_levels),
    #convert to %
    base_vote_share_percent = base_vote_share * 100
  )


```

Process beta (Slopes): create a data frame with only the betas
```{r}
#process beta (slopes)
#4000 obs because 4 chains of 1000 iterations
#15 variables because 5 parties * 3 predictors so each [p, k] combo
df_beta_2019 <- as.data.frame(posterior_samples_2019$beta)

df_beta_long_2019 <- df_beta_2019 %>%
  mutate(iteration = row_number()) %>%
  pivot_longer(cols = -iteration, names_to = "param_name", values_to = "value") %>%
  mutate(
    p_idx = as.integer(gsub("(\\d+)\\.(\\d+)", "\\1", param_name)),
    k_idx = as.integer(gsub("(\\d+)\\.(\\d+)", "\\2", param_name)),
    party = party_levels[p_idx],
    predictor = predictor_cols[k_idx]
  ) %>%
  mutate(param_type = "Slope")

```


###Plots

Plot the distribution of the betas by party and predictor.
```{r}
#fdp
p_beta_fdp_2019 <- ggplot(df_beta_long_2019 %>% filter(party == "fdp"), aes(x = value)) +
  geom_density(alpha = 0.8, adjust = 1, fill = party_colors["fdp"]) +
  facet_wrap( ~ predictor, ncol = 1) +
  labs(
    title = "Posterior Distributions of Coefficients by Predictor: FDP",
    subtitle = "Model trained on 2003-2015 | Effect of predictors on FDP vote shares",
    x = "Coefficient Value",
    y = "Density"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    strip.text = element_text(face = "bold", size = 13, color = "black"),
    strip.background = element_rect(fill = "gray90", color = "white"),
    legend.position = "bottom",
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank(),
    plot.margin = margin(t = 10, r = 10, b = 10, l = 10, unit = "pt")
  )

print(p_beta_fdp_2019)


#cvp
p_beta_cvp_2019 <- ggplot(df_beta_long_2019 %>% filter(party == "cvp"), aes(x = value)) +
  geom_density(alpha = 0.8, adjust = 1, fill = party_colors["cvp"]) +
  facet_wrap( ~ predictor, ncol = 1) +
  labs(
    title = "Posterior Distributions of Coefficients by Predictor: CVP",
    subtitle = "Model trained on 2003-2015 | Effect of predictors on CVP vote shares",
    x = "Coefficient Value",
    y = "Density"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    strip.text = element_text(face = "bold", size = 13, color = "black"),
    strip.background = element_rect(fill = "gray90", color = "white"),
    legend.position = "bottom",
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank(),
    plot.margin = margin(t = 10, r = 10, b = 10, l = 10, unit = "pt")
  )

print(p_beta_cvp_2019)


#sp
p_beta_sp_2019 <- ggplot(df_beta_long_2019 %>% filter(party == "sp"), aes(x = value)) +
  geom_density(alpha = 0.8, adjust = 1, fill = party_colors["sp"]) +
  facet_wrap( ~ predictor, ncol = 1) +
  labs(
    title = "Posterior Distributions of Coefficients by Predictor: SP",
    subtitle = "Model trained on 2003-2015 | Effect of predictors on SP vote shares",
    x = "Coefficient Value",
    y = "Density"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    strip.text = element_text(face = "bold", size = 13, color = "black"),
    strip.background = element_rect(fill = "gray90", color = "white"),
    legend.position = "bottom",
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank(),
    plot.margin = margin(t = 10, r = 10, b = 10, l = 10, unit = "pt")
  )

print(p_beta_sp_2019)


#svp
p_beta_svp_2019 <- ggplot(df_beta_long_2019 %>% filter(party == "svp"), aes(x = value)) +
  geom_density(alpha = 0.8, adjust = 1, fill = party_colors["svp"]) +
  facet_wrap( ~ predictor, ncol = 1) +
  labs(
    title = "Posterior Distributions of Coefficients by Predictor: SVP",
    subtitle = "Model trained on 2003-2015 | Effect of predictors on SVP vote shares",
    x = "Coefficient Value",
    y = "Density"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    strip.text = element_text(face = "bold", size = 13, color = "black"),
    strip.background = element_rect(fill = "gray90", color = "white"),
    legend.position = "bottom",
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank(),
    plot.margin = margin(t = 10, r = 10, b = 10, l = 10, unit = "pt")
  )

print(p_beta_svp_2019)


#other
p_beta_other_2019 <- ggplot(df_beta_long_2019 %>% filter(party == "other"), aes(x = value)) +
  geom_density(alpha = 0.8, adjust = 1, fill = party_colors["other"]) +
  facet_wrap( ~ predictor, ncol = 1) +
  labs(
    title = "Posterior Distributions of Coefficients by Predictor: Other",
    subtitle = "Model trained on 2003-2015 | Effect of predictors on Other vote shares",
    x = "Coefficient Value",
    y = "Density"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    strip.text = element_text(face = "bold", size = 13, color = "black"),
    strip.background = element_rect(fill = "gray90", color = "white"),
    legend.position = "bottom",
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank(),
    plot.margin = margin(t = 10, r = 10, b = 10, l = 10, unit = "pt")
  )

print(p_beta_other_2019)

```

```{r}
#health insurance
df_b_pos_hi_2019 <- df_beta_long_2019 %>%
  filter(predictor == "health_insurance_norm") %>%
  group_by(party) %>%
  do(data.frame(density(.$value)[c("x", "y")])) %>%
  ungroup() %>%
  mutate(party = factor(party, levels = party_levels),
         pos_val = ifelse(x>0, y, 0))

p_b_pos_hi_2019 <- ggplot(df_b_pos_hi_2019, aes(x = x)) +
  geom_line(aes(y=y), color="black", linewidth=0.5)+
  geom_area(aes(y=pos_val, fill=party), alpha=0.6)+
  geom_vline(xintercept = 0, linetype = "dotted", color = "black", linewidth = 0.5) +
  facet_wrap(~ party, ncol = 1, labeller = as_labeller(party_labels)) +
  scale_fill_manual(values = party_colors, guide = "none") +
  scale_y_continuous(breaks = c(0,6), limits = c(0, 8)) +
  scale_x_continuous(breaks = seq(-0.4, 0.4, by = 0.2), limits = c(-0.5, 0.5)) +
  labs(
    x = "Coefficient Value",
    y = "Density"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    strip.text = element_text(face = "bold",size = 13, color = "black"),
    strip.background = element_rect(fill = "gray90", color = "white"),
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_line(linetype = "dashed", linewidth = 0.4),
    plot.margin = margin(t = 10, r = 10, b = 10, l = 10, unit = "pt")
  )

print(p_b_pos_hi_2019)

ggsave("Posterior Distributions of the Health Insurance Coefficient (2019).png", p_b_pos_hi_2019, width = 6, height = 5, dpi = 200, units = "in")

```

```{r}
#health insurance
df_b_hi_2019 <- df_beta_long_2019 %>%
  filter(predictor == "health_insurance_norm") %>%
  mutate(party = factor(party, levels = party_levels))

p_beta_hi_2019 <- ggplot(df_b_hi_2019, aes(x = value, fill = party)) +
  geom_density(alpha = 0.8, adjust = 1) +
  facet_wrap(~ party, ncol = 1) +
  scale_fill_manual(values = party_colors, guide = "none") +
  labs(
    title = "Posterior Distributions of the Health Insurance Coefficient",
    x = "Coefficient Value",
    y = "Density"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    strip.text = element_text(face = "bold",size = 13, color = "black"),
    strip.background = element_rect(fill = "gray90", color = "white"),
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank(),
    plot.margin = margin(t = 10, r = 10, b = 10, l = 10, unit = "pt")
  )

print(p_beta_hi_2019)

```

```{r}
#smi
df_b_pos_smi_2019 <- df_beta_long_2019 %>%
  filter(predictor == "smi_norm") %>%
  group_by(party) %>%
  do(data.frame(density(.$value)[c("x", "y")])) %>%
  ungroup() %>%
  mutate(party = factor(party, levels = party_levels),
         pos_val = ifelse(x>0, y, 0))

p_b_pos_smi_2019 <- ggplot(df_b_pos_smi_2019, aes(x = x)) +
  geom_line(aes(y=y), color="black", linewidth=0.5)+
  geom_area(aes(y=pos_val, fill=party), alpha=0.6)+
  geom_vline(xintercept = 0, linetype = "dotted", color = "black", linewidth = 0.5) +
  facet_wrap(~ party, ncol = 1, labeller = as_labeller(party_labels)) +
  scale_fill_manual(values = party_colors, guide = "none") +
  scale_y_continuous(breaks = c(0,6), limits = c(0, 8)) +
  scale_x_continuous(breaks = seq(-0.4, 0.4, by = 0.2), limits = c(-0.5, 0.5)) +
  labs(
    x = "Coefficient Value",
    y = "Density"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    strip.text = element_text(face = "bold",size = 13, color = "black"),
    strip.background = element_rect(fill = "gray90", color = "white"),
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_line(linetype = "dashed", linewidth = 0.4),
    plot.margin = margin(t = 10, r = 10, b = 10, l = 10, unit = "pt")
  )

print(p_b_pos_smi_2019)

ggsave("Posterior Distributions of the SMI Coefficient (2019).png", p_b_pos_smi_2019, width = 6, height = 5, dpi = 200, units = "in")


```

```{r}
#smi
df_b_smi_2019 <- df_beta_long_2019 %>%
  filter(predictor == "smi_norm") %>%
  mutate(party = factor(party, levels = party_levels))

p_beta_smi_2019 <- ggplot(df_b_smi_2019, aes(x = value, fill = party)) +
  geom_density(alpha = 0.8, adjust = 1) +
  facet_wrap(~ party, ncol = 1) +
  scale_fill_manual(values = party_colors, guide = "none") +
  labs(
    title = "Posterior Distributions of the SMI Coefficient",
    x = "Coefficient Value",
    y = "Density"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    strip.text = element_text(face = "bold",size = 13, color = "black"),
    strip.background = element_rect(fill = "gray90", color = "white"),
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank(),
    plot.margin = margin(t = 10, r = 10, b = 10, l = 10, unit = "pt")
  )

print(p_beta_smi_2019)

```


```{r}
#immigration
df_b_pos_immig_2019 <- df_beta_long_2019 %>%
  filter(predictor == "immigration_norm") %>%
  group_by(party) %>%
  do(data.frame(density(.$value)[c("x", "y")])) %>%
  ungroup() %>%
  mutate(party = factor(party, levels = party_levels),
         pos_val = ifelse(x>0, y, 0))

p_b_pos_immig_2019 <- ggplot(df_b_pos_immig_2019, aes(x = x)) +
  geom_line(aes(y=y), color="black", linewidth=0.5)+
  geom_area(aes(y=pos_val, fill=party), alpha=0.6)+
  geom_vline(xintercept = 0, linetype = "dotted", color = "black", linewidth = 0.5) +
  facet_wrap(~ party, ncol = 1, labeller = as_labeller(party_labels)) +
  scale_fill_manual(values = party_colors, guide = "none") +
  scale_y_continuous(breaks = c(0,6), limits = c(0, 8)) +
  scale_x_continuous(breaks = seq(-0.4, 0.4, by = 0.2), limits = c(-0.5, 0.5)) +
  labs(
    x = "Coefficient Value",
    y = "Density"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    strip.text = element_text(face = "bold",size = 13, color = "black"),
    strip.background = element_rect(fill = "gray90", color = "white"),
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_line(linetype = "dashed", linewidth = 0.4),
    plot.margin = margin(t = 10, r = 10, b = 10, l = 10, unit = "pt")
  )

print(p_b_pos_immig_2019)

ggsave("Posterior Distributions of the Immigration Coefficient (2019).png", p_b_pos_immig_2019, width = 6, height = 5, dpi = 200, units = "in")


```

```{r}
#immigration
df_b_immig_2019 <- df_beta_long_2019 %>%
  filter(predictor == "immigration_norm") %>%
  mutate(party = factor(party, levels = party_levels))

p_beta_immig_2019 <- ggplot(df_b_immig_2019, aes(x = value, fill = party)) +
  geom_density(alpha = 0.8, adjust = 1) +
  facet_wrap(~ party, ncol = 1) +
  scale_fill_manual(values = party_colors, guide = "none") +
  labs(
    title = "Posterior Distributions of the Immigration Coefficient",
    x = "Coefficient Value",
    y = "Density"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    strip.text = element_text(face = "bold",size = 13, color = "black"),
    strip.background = element_rect(fill = "gray90", color = "white"),
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank()
  )

print(p_beta_immig_2019)

```

```{r}
cat("\n--- Posterior Summary (Slopes) ---\n")
df_summary <- df_beta_long_2019 %>%
  group_by(party, predictor) %>%
  summarise(
    mean = mean(value),
    lower = quantile(value, 0.025),
    upper = quantile(value, 0.975),
    significant = (lower > 0) | (upper < 0),
    .groups = 'drop'
  )
print(df_summary)

```


###BETA 0
Plot the beta0s by party.
```{r
plot_beta0s_2019 <- ggplot(df_beta0_long_2019, aes(x = value, fill = party)) +
  geom_density(alpha = 0.7, adjust = 1) +
  facet_wrap(~ party, ncol = 1) +
  scale_fill_manual(values = party_colors, guide = "none") +
  labs(
    title = "Posterior Distribution of Intercepts (b0) for All Parties",
    x = "Intercept Value",
    y = "Density"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    strip.text = element_text(face = "bold", size = 13, color = "black"),
    strip.background = element_rect(fill = "gray90", color = "white"),
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank()
  )

print(plot_beta0s_2019)

```

```{r
plot_beta0_2019 <- ggplot(df_beta0_share_long, aes(x = base_vote_share_percent, fill = party)) +
  geom_density(alpha = 0.7, adjust = 1) +
  facet_wrap(~ party, ncol = 1, scales = "free_x") +
  scale_fill_manual(values = party_colors, guide = "none") +
  labs(
    title = "Model-Implied Base Support (Intercepts b0)",
    subtitle = "Predicted Vote Share if all Predictors were Average",
    x = "Base Vote Share (%)",
    y = "Density"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    strip.text = element_text(face = "bold", size = 13, color = "black"),
    strip.background = element_rect(fill = "gray90", color = "white"),
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank()
  )

print(plot_beta0_2019)

```



```{r}
plot_beta0_violin_2019 <- ggplot(df_beta0_share_long_2019, aes(x = party, y = base_vote_share_percent, fill = party)) +
  geom_violin(alpha = 0.8, scale = "width", trim = FALSE, width = 0.6) +
  geom_boxplot(width = 0.1, color = "white", alpha = 0.5, outlier.shape = NA) +
  scale_fill_manual(values = party_colors, guide = "none") +
  scale_x_discrete(labels = party_labels) +
  labs(
    title = "Distribution of Base Support by Parties (2019)",
    x = "Party",
    y = "Base Vote Share (%)"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(face = "bold", size = 12, color = "black"),
    panel.grid.minor = element_blank(),
    plot.margin = margin(t = 10, r = 10, b = 10, l = 10, unit = "pt")
  )

print(plot_beta0_violin_2019)

```



```{r}
cat("\n--- Summary of Base Support Distributions ---\n")
cat("Values represent Predicted Vote Share (%) if Predictors = 0\n\n")

summary_b0_2019 <- df_beta0_share_long_2019 %>%
  group_by(party) %>%
  summarise(
    mean_share = mean(base_vote_share_percent),
    median_share = median(base_vote_share_percent),
    lower_ci = quantile(base_vote_share_percent, 0.025),
    upper_ci = quantile(base_vote_share_percent, 0.975),
    min = min(base_vote_share_percent),
    max = max(base_vote_share_percent),
    .groups = "drop"
  ) %>%
  mutate(party = factor(party, levels = party_levels)) %>%
  arrange(party)

print(summary_b0_2019)
```

```{r}
p_b0_2019 <- ggplot(summary_b0_2019, aes(x = party)) +
  #actual result
  geom_rect(data = df_results_2019,
            aes(width = 0.4,
                ymin = 0, ymax = vote_share, fill = party, alpha = 0.6),
            color = NA, show.legend = FALSE) +
  #95% CI
  geom_rect(aes(width = 0.3,
                ymin = lower_ci / 100, ymax = upper_ci / 100),
            fill = "gray70", alpha = 0.6, color = NA) +
  #mean
  geom_point(aes(y = mean_share / 100),
             shape = 21, fill = "white", color = "black", size = 3, stroke = 0.5) +
  #scale y to percentage
  scale_y_continuous(labels = scales::percent_format(accuracy = 1),
                     limits = c(0, 0.4),
                     expand = expansion(mult = c(0, 0.1))) +
  #party colors
  scale_fill_manual(values = party_colors) +
  #values next to bars
  geom_text(aes(y = mean_share / 100, label = sprintf("%.1f", mean_share)),
            hjust = -0.9, size = 4.5, color = "gray40") +
  geom_text(data = df_results_2019, aes(y = vote_share, label = sprintf("%.1f", vote_share * 100)),
            hjust = 1.9, size = 4.5, color = "black") +
  #party names lowercase to capital
  scale_x_discrete(labels = party_labels) +
  labs(
    x = "Party",
    y = "Vote Share"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_line(color = "gray90", linetype = "dashed"),
    panel.grid.major.x = element_blank(),
    axis.text.x = element_text(face = "bold", size = 12),
    axis.text.y = element_text(size = 12),
    axis.ticks.x = element_blank(),
    plot.margin = margin(t = 10, r = 10, b = 10, l = 10, unit = "pt")
  )

print(p_b0_2019)

ggsave("Distribution of Base Support by Parties (2019).png", p_b0_2019, width = 8, height = 5, dpi = 200, units = "in")

```













