---
title: "predictors & run fundamentals_PPC"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Predictors
(The data points are from the years 2011 to 2024.)
Since we only care about election years, we only consider the data points from 2003, 2007, 2011, 2015, 2019 and 2023.
(Immigration data collection started in 2011, so the values 0 are to be interpreted as NA.)
SMI % return is over 4 years - we compare the values from last election year to current election year. Similarly the % increase of health insurance premiums is computed.
I center all values, which is especially important for immigration as then the value 0 then implies "average immigration" for the election years 2003 and 2007.

```{r}
#parties and predictors
party_levels <- c("fdp", "cvp", "sp", "svp", "other")
predictor_cols <- c("smi_norm", "health_insurance_norm", "immigration_norm")

party_colors <- c(
  "fdp" = "cornflowerblue",
  "cvp" = "darkorange",
  "sp" = "red2",
  "svp" = "green4",
  "other" = "plum2"
)

party_labels <- c(
  "fdp" = "FDP",
  "cvp" = "CVP",
  "sp" = "SP",
  "svp" = "SVP",
  "other" = "Other"
)


```


```{r}
library(readr)

election_data <- read_csv("swiss election data.csv")

```


```{r}
#predictor data from 2003-2023 so we will need to only extract the data points from those years from the election_data

election_years <- c(2003, 2007, 2011, 2015, 2019, 2023)

```


```{r}
#CODE TO RUN IN CONSOLE TO RUN MODEL

library(rstan)
library(dplyr)

setwd("/Users/juliahollmann/Documents/GitHub/master-thesis/code")

set.seed(266)

#looking at 6 elections and 5 parties (4 main + rest grouped as 1 "other")
N_e <- length(election_years)
N_p <- 5
N_obs <- N_e * N_p

df <- election_data %>%
  filter(year %in% election_years) %>%
  arrange(year, party) %>%
  mutate(
    party = factor(party, levels = c("fdp", "cvp", "sp", "svp", "other")),
    e_id_temp = factor(year, levels = election_years)
  )

X_matrix <- as.matrix(df[, c("smi_norm", "health_insurance_norm", "immigration_norm")])

p_id <- as.integer(df$party)
e_id <- as.integer(df$e_id_temp)


```

```{r}
stan_data <- list(
  N_e = N_e,
  N_p = N_p,
  N_obs = N_obs,
  K = 3,
  X = X_matrix,
  p_id = p_id,
  e_id = e_id
)

```


```{r}
fit <- stan(
  file = "fundamentals_PPC.stan",
  data = stan_data,
  iter = 2000,
  chains = 4,
  refresh = 500
)

print(fit, pars = c("beta_0", "beta"))

saveRDS(fit, "fundamentals_fit_ppc.rds")

```


```{r}
library(ggplot2)
library(dplyr)
library(scales)

#readRDS("fundamentals_fit_ppc.rds")

sims <- rstan::extract(fit, "v_sim")$v_sim

#aggregate all simulations into single vector so can create comparable plot to nationalrat_vote_share_histogram
all_sim_shares <- as.vector(sims)

#create data frame
df_sim <- data.frame(vote_share = all_sim_shares)

#stats
avg_sim_share <- mean(all_sim_shares)
min_sim_share <- min(all_sim_shares)
max_sim_share <- max(all_sim_shares)
median_sim_share <- median(all_sim_shares)

bin_width <- 0.01
#create breaks explicitly
breaks <- seq(0, max_sim_share + bin_width, by = bin_width)
#cut data into bins
bins <- cut(all_sim_shares, breaks = breaks, include.lowest = TRUE, right = TRUE)
#count frequencies
bin_counts <- table(bins)
#find index of the highest bin
highest_bin_index <- which.max(bin_counts)
#calculate center directly from breaks vector
lower_bound <- breaks[highest_bin_index]
upper_bound <- breaks[highest_bin_index + 1]
modal_class_center <- (lower_bound + upper_bound) / 2

#histogram
p_sim <- ggplot(df_sim, aes(x=vote_share)) +
  geom_histogram(binwidth = bin_width, boundary = 0, color = "black", fill = "lightblue") +
  geom_vline(xintercept = avg_sim_share, color = "red", linewidth = 1, linetype = "solid") +
  geom_vline(xintercept = c(min_sim_share, max_sim_share), color = "darkgreen", linewidth = 1, linetype = "dashed") +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  labs(
    title = "Distribution of Simulated Vote Shares",
    subtitle = "Bin width: 1% (0.01) | All Parties and Elections Combined",
    caption = "Red: Simulated Mean | Green: Simulated Minimum / Maximum",
    x = "Vote Share",
    y = "Frequency"
  ) +
  theme_minimal() +
  theme(
    margins = margin(t = 10, r = 10, b = 10, l = 10, unit = "pt")
  )

print(p_sim)


#density curve
p_sim_curve <- ggplot(df_sim, aes(x=vote_share))+
  geom_density(fill="lightblue", color="black", alpha=0.6, linewidth=1)+
  #historical mean
  geom_vline(xintercept = mean(election_data$vote_share), color="red", linewidth=1)+
  #historical min
  geom_vline(xintercept = min(election_data$vote_share), color="darkgreen", linewidth=1, linetype="dashed")+
  #historical max
  geom_vline(xintercept = max(election_data$vote_share), color="darkgreen", linewidth=1, linetype="dashed")+
  scale_x_continuous(labels=scales::percent_format(accuracy=1))+
  scale_y_continuous(expand = expansion(mult=c(0,0.05)))+
  labs(
    x = "Vote Share",
    y = "Density"
  )+
  theme_minimal() +
  theme(
    margins = margin(t = 10, r = 10, b = 10, l = 10, unit = "pt")
  )

print(p_sim_curve)

ggsave("Distribution of Simulated Vote Shares PPC (Bin Width 0.01).png", p_sim_curve, width = 6, height = 4, dpi = 200, units = "in")

#rug plot
p_rug <- p_sim_curve + geom_rug(
  data = election_data,
  aes(x = vote_share),
  sides = "b"
)

print(p_rug)

ggsave("Distribution of Simulated Vote Shares PPC (Bin Width 0.01) Rug Plot.png", p_rug, width = 6, height = 4, dpi = 200, units = "in")

#summary
cat("--Simulated Statistics--\n",
    "Minimum Vote Share: ", min_sim_share,
    "\nMaximum Vote Share: ", max_sim_share,
    "\nAverage Vote Share: ", avg_sim_share,
    "\nMedian Vote Share: ", median_sim_share,
    "\nModal Bin Center: ", modal_class_center)

```


```{r}
hist(all_sim_shares,
     col = "lightblue",
     border = "white",
     breaks = 84,
     xlim = c(0, 0.35),
     main = "Simulated Vote Shares")

abline(v = avg_sim_share, col = "red", lwd = 2)
abline(v = median_sim_share, col = "purple", lwd = 2)
abline(v = modal_class_center, col = "orange", lwd = 2)

abline(v = min_sim_share, col = "green", lwd = 2)
abline(v = max_sim_share, col = "green", lwd = 2)
```


###Simulation DF Detailed

```{r}
library(dplyr)
library(rstan)
library(tidyr)
library(ggplot2)
library(scales)

#extract simulations
sims_array <- rstan::extract(fit, "v_sim")$v_sim
n_iterations <- dim(sims_array)[1]

#create mapping frame for the 30 observations used in the model using df from above
df_model_obs <- df %>%
  arrange(year, party) %>%
  mutate(
    party = factor(party, levels = c("fdp", "cvp", "sp", "svp", "other")),
    year = as.integer(as.character(year))
  )

#verify row count matches N_obs
if(nrow(df_model_obs) != (N_e * N_p)) {
  stop("Mismatch between model observations and mapping dataframe.")
}

#expand to long format for all iterations (repeat the mapping frame for every iteration)
df_sims_long <- data.frame(
  iteration = rep(1:n_iterations, each = nrow(df_model_obs)),
  party = rep(df_model_obs$party, times = n_iterations),
  year = rep(df_model_obs$year, times = n_iterations),
  vote_share = as.vector(t(sims_array))
)

#load full historical data for comparison
df_history <- election_data %>%
  mutate(
    party = factor(party, levels = c("fdp", "cvp", "sp", "svp", "other")),
    year = as.integer(year)
  )

```


####Winners
Check whether or not the same party always wins.

```{r}
#reshape array to [iterations, years, parties]
sims_3d <- array(sims_array, dim = c(n_iterations, N_e, N_p))

#identify winner index (1-5) for each iteration and year
winners_idx <- apply(sims_3d, c(1,2), which.max)

#map indices to party names
party_levels <- c("fdp", "cvp", "sp", "svp", "other")
winners_df <- as.data.frame(winners_idx)
colnames(winners_df) <- election_years

#convert to long format for plotting
winners_long <- winners_df %>%
  mutate(iteration = row_number()) %>%
  pivot_longer(cols = -iteration, names_to = "year", values_to = "winner_idx") %>%
  mutate(
    year = as.integer(year),
    winner_party = factor(party_levels[winners_idx], levels = c("fdp", "cvp", "sp", "svp", "other"))
  )

#plot distribution of winners
p_winners <- ggplot(winners_long, aes(x = winner_party, fill = winner_party)) +
  geom_bar(show.legend = FALSE) +
  scale_fill_manual(values = c(
    "fdp" = "cornflowerblue",
    "cvp" = "darkorange",
    "sp" = "red2",
    "svp" = "green4",
    "other" = "plum2"
  )) +
  labs(
    title = "Frequency of Winning Party in Prior Predictive Simulations",
    x = "Winning Party",
    y = "Count"
  ) +
  theme_minimal()

print(p_winners)

#quick console check: proportion of wins
prop_table <- prop.table(table(winners_long$winner_party))
print("Proportion of wins per party:")
print(prop_table)

```


####Histograms per party over time

```{r}
p_facet <- ggplot(df_sims_long, aes(x = vote_share)) +
  geom_histogram(binwidth = 0.01, color = "black", alpha = 0.8) +
  facet_grid(party ~ year) + #rows: parties, columns: years
  scale_x_continuous(labels = percent_format(Accuracy = 1), limits = c(0, 0.5)) +
  labs(
    title = "Prior Predictive Check: Vote Share Distributions",
    x = "Simulated Vote Share",
    y = "Frequency"
  ) +
  theme_minimal() +
  theme(
    strip.text = element_text(size = 8, face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

print(p_facet)

#overlay historical mean for context
hist_results <- df_history %>%
  filter(year %in% election_years) %>%
  select(party, year, vote_share)

p_facet_overlay <- p_facet +
  geom_vline(data = hist_results, aes(xintercept = vote_share), color = "red", linewidth = 0.8, linetype = "dashed") +
  labs(subtitle = "Red dashed line = Historical Vote Share for that Year/Party")

print(p_facet_overlay)

```


###95% CI

```{r}
df_sims_stats <- df_sims_long %>%
  group_by(party) %>%
  summarise(
    mean_sim = mean(vote_share),
    sd_sim = sd(vote_share),
    lower_ci = quantile(vote_share, 0.025),
    upper_ci = quantile(vote_share, 0.975),
    .groups = "drop"
  )

p_sims_ci <- ggplot(df_sims_stats, aes(x = party, y = mean_sim)) +
  #95% CI
  geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci),
                width = 0.2,
                color = "gray40",
                linewidth = 1) +
  #mean prediction point
  geom_point(size = 4, color = "gray30", shape = 21, fill ="white") +
  #scale y to percentage
  scale_y_continuous(labels = scales::percent_format(accuracy = 1),
                     limits = c(0, NA),
                     expand = expansion(mult = c(0, 0.1))) +
  labs(
    title = "Prior Predictive Check - 95% CI",
    x = "Party",
    y = "Vote Share"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(face = "bold", color = party_colors),
    margins = margin(t = 10, r = 10, b = 10, l = 10, unit = "pt")
  )

print(p_sims_ci)

#ggsave("Prior Predictive Check 95pct CI.png", p_sims_ci, width = 6, height = 4, dpi = 200, units = "in")

```

```{r}

cat("---simulation stats---\n")
print(df_sims_stats)

```


```{r}
p_sims_ci_new <- ggplot(df_sims_stats, aes(x = party)) +
  #95% CI
  geom_rect(aes(width = 0.3,
                ymin = lower_ci, ymax = upper_ci),
            fill = "gray50", alpha = 0.6, color = NA) +
  #mean
  geom_point(aes(y = mean_sim),
             shape = 21, fill = "white", color = "black", size = 3, stroke = 0.5) +
  #scale y to percentage
  scale_y_continuous(labels = scales::percent_format(accuracy = 1),
                     limits = c(0, 0.4),
                     expand = expansion(mult = c(0, 0.1))) +
  #values next to bars
  geom_text(aes(y = mean_sim, label = sprintf("%.1f", mean_sim * 100)),
            hjust = -0.9, size = 4.5, color = "gray40") +
  #party names lowercase to capital
  scale_x_discrete(labels = party_labels) +
  labs(
    x = "Party",
    y = "Vote Share"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_line(color = "gray90", linetype = "dashed"),
    panel.grid.major.x = element_blank(),
    axis.text.x = element_text(face = "bold", size = 12),
    axis.text.y = element_text(size = 12),
    axis.ticks.x = element_blank(),
    plot.margin = margin(t = 20, r = 20, b = 10, l = 10, unit = "pt")
  )

print(p_sims_ci_new)

ggsave("Prior Predictive Check 95pct CI.png", p_sims_ci_new, width = 8, height = 5, dpi = 200, units = "in")


```
















