---
title: "swiss_polling_data_summary"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#setup
library(ggplot2)
library(dplyr)
library(scales)
library(lubridate)
library(zoo)
library(pracma)
library(tidyverse)

```

#Swiss Polling Data Summary
Import the csv file *swiss polling data* and create plots that show how the vote shares change over time (waves) by election year and by company, and compute the average vote shares over all companies by time periods and plot the change in vote shares over time.

```{r}
#import swiss polling data csv

library(readr)

swiss_polling_data <- read_csv("swiss polling data.csv")

#convert polling_period strings to date
swiss_polling_data$polling_period <- as.Date(swiss_polling_data$polling_period, format = "%d.%m.%Y")

#lowercase to uppercase
company_labels <- c(
  "fors" = "FORS",
  "sotomo" = "Sotomo",
  "tamedia" = "Tamedia"
)

```

Import the csv file *swiss election data* that includes the election results from 1971 to 2023.

```{r}
#import election results 2019 and 2023
election_data <- read.csv("swiss election data.csv")

results_2019 <- election_data %>%
  filter(year == 2019) %>%
  mutate(vote_share = vote_share *100)

results_2023 <- election_data %>%
  filter(year == 2023) %>%
  mutate(vote_share = vote_share *100)

```

Import the csv files *fundamentals predictions* for 2019 and 2023 that include the fundamentals predictions for both years.

```{r}
#2019
predictions_2019 <- read_csv("fundamentals_predictions_2019.csv")
predictions_2019 <- predictions_2019 %>%
  mutate(mean_pred = mean_pred *100)

#fdp
fdp_pred_2019 <- predictions_2019 %>%
  filter(party == "fdp")

#cvp
cvp_pred_2019 <- predictions_2019 %>%
  filter(party == "cvp")

#sp
sp_pred_2019 <- predictions_2019 %>%
  filter(party == "sp")

#svp
svp_pred_2019 <- predictions_2019 %>%
  filter(party == "svp")

#other
other_pred_2019 <- predictions_2019 %>%
  filter(party == "other")


#2023
predictions_2023 <- read_csv("fundamentals_predictions_2023.csv")
predictions_2023 <- predictions_2023 %>%
  mutate(mean_pred = mean_pred *100)

#fdp
fdp_pred_2023 <- predictions_2023 %>%
  filter(party == "fdp")

#cvp
cvp_pred_2023 <- predictions_2023 %>%
  filter(party == "cvp")

#sp
sp_pred_2023 <- predictions_2023 %>%
  filter(party == "sp")

#svp
svp_pred_2023 <- predictions_2023 %>%
  filter(party == "svp")

#other
other_pred_2023 <- predictions_2023 %>%
  filter(party == "other")

```

##Setup

```{r}
#assign colors and names to parties
party_colors <- c(
  "fdp" = "cornflowerblue",
  "cvp" = "darkorange",
  "sp" = "red2",
  "svp" = "green4",
  "other" = "plum2"
)

party_labels <- c(
  "fdp" = "FDP",
  "cvp" = "CVP",
  "sp" = "SP",
  "svp" = "SVP",
  "other" = "Other"
)

#filter polling data by year
SPD_2019 <- swiss_polling_data %>%
  filter(year == 2019)

SPD_2023 <- swiss_polling_data %>%
  filter(year == 2023)

```

####Filter Data by Polling Company / Party and Year

```{r}
#tamedia
SPD_tamedia <- swiss_polling_data %>%
  filter(polling_company == "tamedia")

SPD_tamedia_2019 <- SPD_tamedia %>%
  filter(year == 2019)

SPD_tamedia_2023 <- SPD_tamedia %>%
  filter(year == 2023)

```

```{r}
#sotomo
SPD_sotomo <- swiss_polling_data %>%
  filter(polling_company == "sotomo")

SPD_sotomo_2019 <- SPD_sotomo %>%
  filter(year == 2019)

SPD_sotomo_2023 <- SPD_sotomo %>%
  filter(year == 2023)

```

```{r}
#fors
SPD_fors <- swiss_polling_data %>%
  filter(polling_company == "fors")

SPD_fors_2019 <- SPD_fors %>%
  filter(year == 2019)

SPD_fors_2023 <- SPD_fors %>%
  filter(year == 2023)

```

```{r}
#fdp
fdp_polls <- swiss_polling_data %>%
  filter(party == "fdp") %>%
  arrange(polling_period)

fdp_polls_2019 <- fdp_polls %>%
  filter(year == 2019)

fdp_polls_2023 <- fdp_polls %>%
  filter(year == 2023)


#cvp
cvp_polls <- swiss_polling_data %>%
  filter(party == "cvp") %>%
  arrange(polling_period)

cvp_polls_2019 <- cvp_polls %>%
  filter(year == 2019)

cvp_polls_2023 <- cvp_polls %>%
  filter(year == 2023)


#sp
sp_polls <- swiss_polling_data %>%
  filter(party == "sp") %>%
  arrange(polling_period)

sp_polls_2019 <- sp_polls %>%
  filter(year == 2019)

sp_polls_2023 <- sp_polls %>%
  filter(year == 2023)


#svp
svp_polls <- swiss_polling_data %>%
  filter(party == "svp") %>%
  arrange(polling_period)

svp_polls_2019 <- svp_polls %>%
  filter(year == 2019)

svp_polls_2023 <- svp_polls %>%
  filter(year == 2023)


#other
other_polls <- swiss_polling_data %>%
  filter(party == "other") %>%
  arrange(polling_period)

other_polls_2019 <- other_polls %>%
  filter(year == 2019)

other_polls_2023 <- other_polls %>%
  filter(year == 2023)

```

###Functions

```{r}
#aggregate same day polls into one weighted average
#this way get one ewma value per day later

aggregate_polls <- function(data) {
  data %>%
    group_by(party, polling_period) %>%
    summarise(
      poll_vote_share = sum(poll_vote_share * sample_size, na.rm = TRUE) / sum(sample_size, na.rm = TRUE),
      sample_size = sum(sample_size, na.rm = TRUE),
      .groups = "drop"
    )
}
```


```{r}
#EWMA function
calc_ewma <- function(votes, samples, l_base = 0.5) {
  n <-  length(votes)
  ewma  <-  numeric(n)
  
  ewma[1] <- votes[1]
  
  weights <- samples / max(samples, na.rm = TRUE)
  
  for (i in 2:n) {
    current_l <- l_base * weights[i]
    ewma[i] <- (current_l * votes[i]) + (1 - current_l) * ewma[i-1]
  }
  
  return(ewma)
}

```

##Plots comparing trajectories of public opinion by polling company

```{r}
#plot 2019 all three companies stacked
plot_2019 <- ggplot(SPD_2019, aes(x = polling_period, y = poll_vote_share, color = party))+
  geom_line(linewidth = 0.5)+
  geom_point(size = 1)+
  facet_grid(. ~ polling_company, scales = "fixed", labeller = as_labeller(company_labels))+
  scale_color_manual(name = "", values = party_colors, labels = party_labels)+
  scale_y_continuous(limits = c(5, 40), breaks = seq(5, 40, by = 5), minor_breaks = seq(5, 40, by = 2.5), expand = expansion(0))+
  labs(
    x = "Polling Period",
    y = "Vote Share (%)"
  )+
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.x = element_text(margin = margin(t = 10, unit = "pt")),
    legend.position = "bottom",
    strip.background = element_rect(fill = "gray90", color = "gray9"),
    panel.background = element_rect(fill = "gray100", color = "gray9"),
    plot.margin = margin(t = 20, r = 20, b = 0, l = 10, unit = "pt"))

print(plot_2019)

ggsave("Polls by Company (2019).png", plot_2019, width = 8, height = 5, dpi = 200, units = "in")

```

```{r}
#plot 2023 all three companies stacked
plot_2023 <- ggplot(SPD_2023, aes(x = polling_period, y = poll_vote_share, color = party))+
  geom_line(linewidth = 0.5)+
  geom_point(size = 1)+
  facet_grid(. ~ polling_company, scales = "fixed", labeller = as_labeller(company_labels))+
  scale_color_manual(name = "", values = party_colors, labels = party_labels)+
  scale_y_continuous(limits = c(5, 40), breaks = seq(5, 40, by = 5), minor_breaks = seq(5, 40, by = 2.5), expand = expansion(0))+
  labs(
    x = "Polling Period",
    y = "Vote Share (%)"
  )+
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.x = element_text(margin = margin(t = 10, unit = "pt")),
    legend.position = "bottom",
    strip.background = element_rect(fill = "gray90", color = "gray9"),
    panel.background = element_rect(fill = "gray100", color = "gray9"),
    plot.margin = margin(t = 20, r = 20, b = 0, l = 10, unit = "pt"))

print(plot_2023)

ggsave("Polls by Company (2023).png", plot_2023, width = 8, height = 5, dpi = 200, units = "in")

```

##Plots by Party
Plots of each party's public opinion trajectory - combining all polls.
Exponential weighted moving average (EWMA) to compute the average public opinion from all polls.

###Setup

```{r}
ewma_2019 <- SPD_2019 %>%
  aggregate_polls() %>%
  arrange(party, polling_period) %>%
  group_by(party) %>%
  mutate(
    ewma_value = calc_ewma(poll_vote_share, sample_size, a_base = 0.5)
  ) %>%
  ungroup() %>%
  select(polling_period, party, ewma_value) %>%
  pivot_wider(
    id_cols = polling_period,
    names_from = party,
    values_from = ewma_value,
    names_prefix = "ewma_"
  ) %>%
  arrange(polling_period)

write_csv(ewma_2019, "ewma_2019.csv")

```

```{r}
ewma_2023 <- SPD_2023 %>%
  aggregate_polls() %>%
  arrange(party, polling_period) %>%
  group_by(party) %>%
  mutate(
    ewma_value = calc_ewma(poll_vote_share, sample_size, a_base = 0.5)
  ) %>%
  ungroup() %>%
  select(polling_period, party, ewma_value) %>%
  pivot_wider(
    id_cols = polling_period,
    names_from = party,
    values_from = ewma_value,
    names_prefix = "ewma_"
  ) %>%
  arrange(polling_period)

write_csv(ewma_2023, "ewma_2023.csv")

```

```{r}
#election result fdp 2019
fdp_result_2019 <- results_2019$vote_share[1]

#EWMA (exponential weighted moving average)
fdp_ewma_2019 <- ewma_2019 %>%
  select(polling_period, ewma_fdp) %>%
  rename(ewma = ewma_fdp)

```

```{r}
#election result fdp 2023
fdp_result_2023 <- results_2023$vote_share[1]

#EWMA (exponential weighted moving average)
fdp_ewma_2023 <- ewma_2023 %>%
  select(polling_period, ewma_fdp) %>%
  rename(ewma = ewma_fdp)

```

```{r}
#election result cvp 2019
cvp_result_2019 <- results_2019$vote_share[2]

#EWMA (exponential weighted moving average)
cvp_ewma_2019 <- ewma_2019 %>%
  select(polling_period, ewma_cvp) %>%
  rename(ewma = ewma_cvp)

```

```{r}
#election result cvp 2023
cvp_result_2023 <- results_2023$vote_share[2]

#EWMA (exponential weighted moving average)
cvp_ewma_2023 <- ewma_2023 %>%
  select(polling_period, ewma_cvp) %>%
  rename(ewma = ewma_cvp)

```

```{r}
#election result sp 2019
sp_result_2019 <- results_2019$vote_share[3]

#EWMA (exponential weighted moving average)
sp_ewma_2019 <- ewma_2019 %>%
  select(polling_period, ewma_sp) %>%
  rename(ewma = ewma_sp)

```

```{r}
#election result sp 2023
sp_result_2023 <- results_2023$vote_share[3]

#EWMA (exponential weighted moving average)
sp_ewma_2023 <- ewma_2023 %>%
  select(polling_period, ewma_sp) %>%
  rename(ewma = ewma_sp)

```

```{r}
#election result svp 2019
svp_result_2019 <- results_2019$vote_share[4]

#EWMA (exponential weighted moving average)
svp_ewma_2019 <- ewma_2019 %>%
  select(polling_period, ewma_svp) %>%
  rename(ewma = ewma_svp)

```

```{r}
#election result svp 2023
svp_result_2023 <- results_2023$vote_share[4]

#EWMA (exponential weighted moving average)
svp_ewma_2023 <- ewma_2023 %>%
  select(polling_period, ewma_svp) %>%
  rename(ewma = ewma_svp)

```

```{r}
#election result other 2019
other_result_2019 <- results_2019$vote_share[5]

#EWMA (exponential weighted moving average)
other_ewma_2019 <- ewma_2019 %>%
  select(polling_period, ewma_other) %>%
  rename(ewma = ewma_other)

```

```{r}
#election result other 2023
other_result_2023 <- results_2023$vote_share[5]

#EWMA (exponential weighted moving average)
other_ewma_2023 <- ewma_2023 %>%
  select(polling_period, ewma_other) %>%
  rename(ewma = ewma_other)

```

###FDP

```{r}
#plot 2019
plot_all_fdp_2019 <- ggplot() +
  geom_point(data = fdp_polls_2019,
             aes(x = polling_period, y = poll_vote_share, shape = factor(polling_company)),
             size = 2) +
  geom_hline(yintercept = fdp_result_2019, color = "cornflowerblue", linewidth = 1) +
  geom_hline(yintercept = fdp_pred_2019$mean_pred, linetype = "dashed", linewidth = 0.4) +
  geom_rect(data = fdp_pred_2019,
            aes(xmin = -Inf, xmax = Inf, ymin = lower_ci *100, ymax = upper_ci *100),
            fill = "gray", alpha = 0.2, inherit.aes = FALSE) +
  geom_line(data = fdp_ewma_2019,
            aes(x = polling_period, y = ewma),
            color = "black") +
  scale_y_continuous(limits = c(5, 30), breaks = seq(5, 30, by = 5), minor_breaks = seq(5, 30, by = 2.5), expand = expansion(0)) +
  labs(
    x = "Polling Period",
    y = "Vote Share (%)",
    shape = "Polling Company"
  ) +
  scale_shape_discrete(name = "", labels = company_labels) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    strip.background = element_rect(color = "white"),
    plot.margin = margin(t = 20, r = 20, b = 0, l = 10, unit = "pt")
  )


print(plot_all_fdp_2019)

ggsave("FDP Pre-Election Polling Outcomes Compared to the Election Result (2019).png", plot_all_fdp_2019, width = 6, height = 4, dpi = 200, units = "in")

```

```{r}
#plot 2023
plot_all_fdp_2023 <- ggplot() +
  geom_point(data = fdp_polls_2023,
             aes(x = polling_period, y = poll_vote_share, shape = factor(polling_company)),
             size = 2) +
  geom_hline(yintercept = fdp_result_2023, color = "cornflowerblue", linewidth = 1) +
  geom_hline(yintercept = fdp_pred_2023$mean_pred, linetype = "dashed", linewidth = 0.4) +
    geom_rect(data = fdp_pred_2019,
            aes(xmin = -Inf, xmax = Inf, ymin = lower_ci *100, ymax = upper_ci *100),
            fill = "gray", alpha = 0.2, inherit.aes = FALSE) +
  geom_line(data = fdp_ewma_2023,
            aes(x = polling_period, y = ewma),
            color = "black") +
  scale_y_continuous(limits = c(5, 30), breaks = seq(5, 30, by = 5), minor_breaks = seq(5, 30, by = 2.5), expand = expansion(0)) +
  labs(
    x = "Polling Period",
    y = "Vote Share (%)",
    shape = "Polling Company"
  ) +
  scale_shape_discrete(name = "", labels = company_labels) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    strip.background = element_rect(color = "white"),
    plot.margin = margin(t = 20, r = 20, b = 0, l = 10, unit = "pt")
  )


print(plot_all_fdp_2023)

ggsave("FDP Pre-Election Polling Outcomes Compared to the Election Result (2023).png", plot_all_fdp_2023, width = 6, height = 4, dpi = 200, units = "in")

```

###CVP

```{r}
#plot 2019
plot_all_cvp_2019 <- ggplot() +
  geom_point(data = cvp_polls_2019,
             aes(x = polling_period, y = poll_vote_share, shape = factor(polling_company)),
             size = 2) +
  geom_hline(yintercept = cvp_result_2019, color = "darkorange", linewidth = 1) +
  geom_hline(yintercept = cvp_pred_2019$mean_pred, linetype = "dashed", linewidth = 0.4) +
  geom_rect(data = cvp_pred_2019,
            aes(xmin = -Inf, xmax = Inf, ymin = lower_ci *100, ymax = upper_ci *100),
            fill = "gray", alpha = 0.2, inherit.aes = FALSE) +
  geom_line(data = cvp_ewma_2019,
            aes(x = polling_period, y = ewma),
            color = "black") +
  scale_y_continuous(limits = c(0, 25), breaks = seq(0, 25, by = 5), minor_breaks = seq(0, 25, by = 2.5), expand = expansion(0)) +
  labs(
    x = "Polling Period",
    y = "Vote Share (%)",
    shape = "Polling Company"
  ) +
  scale_shape_discrete(name = "", labels = company_labels) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    strip.background = element_rect(color = "white"),
    plot.margin = margin(t = 20, r = 20, b = 0, l = 10, unit = "pt")
  )


print(plot_all_cvp_2019)

ggsave("CVP Pre-Election Polling Outcomes Compared to the Election Result (2019).png", plot_all_cvp_2019, width = 6, height = 4, dpi = 200, units = "in")


```

```{r}
#plot 2023
plot_all_cvp_2023 <- ggplot() +
  geom_point(data = cvp_polls_2023,
             aes(x = polling_period, y = poll_vote_share, shape = factor(polling_company)),
             size = 2) +
  geom_hline(yintercept = cvp_result_2023, color = "darkorange", linewidth = 1) +
  geom_hline(yintercept = cvp_pred_2023$mean_pred, linetype = "dashed", linewidth = 0.4) +
  geom_rect(data = cvp_pred_2023,
            aes(xmin = -Inf, xmax = Inf, ymin = lower_ci *100, ymax = upper_ci *100),
            fill = "gray", alpha = 0.2, inherit.aes = FALSE) +
  geom_line(data = cvp_ewma_2023,
            aes(x = polling_period, y = ewma),
            color = "black") +
  scale_y_continuous(limits = c(0, 25), breaks = seq(0, 25, by = 5), minor_breaks = seq(0, 25, by = 2.5), expand = expansion(0)) +
  labs(
    x = "Polling Period",
    y = "Vote Share (%)",
    shape = "Polling Company"
  ) +
  scale_shape_discrete(name = "", labels = company_labels) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    strip.background = element_rect(color = "white"),
    plot.margin = margin(t = 20, r = 20, b = 0, l = 10, unit = "pt")
  )


print(plot_all_cvp_2023)

ggsave("CVP Pre-Election Polling Outcomes Compared to the Election Result (2023).png", plot_all_cvp_2023, width = 6, height = 4, dpi = 200, units = "in")


```

###SP

```{r}
#plot 2019
plot_all_sp_2019 <- ggplot() +
  geom_point(data = sp_polls_2019,
             aes(x = polling_period, y = poll_vote_share, shape = factor(polling_company)),
             size = 2) +
  geom_hline(yintercept = sp_result_2019, color = "red2", linewidth = 1) +
  geom_hline(yintercept = sp_pred_2019$mean_pred, linetype = "dashed", linewidth = 0.4) +
  geom_rect(data = sp_pred_2019,
            aes(xmin = -Inf, xmax = Inf, ymin = lower_ci *100, ymax = upper_ci *100),
            fill = "gray", alpha = 0.2, inherit.aes = FALSE) +
  geom_line(data = sp_ewma_2019,
            aes(x = polling_period, y = ewma),
            color = "black") +
  scale_y_continuous(limits = c(10, 35), breaks = seq(10, 35, by = 5), minor_breaks = seq(10, 35, by = 2.5), expand = expansion(0)) +
  labs(
    x = "Polling Period",
    y = "Vote Share (%)",
    shape = "Polling Company"
  ) +
  scale_shape_discrete(name = "", labels = company_labels) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    strip.background = element_rect(color = "white"),
    plot.margin = margin(t = 20, r = 20, b = 0, l = 10, unit = "pt")
  )


print(plot_all_sp_2019)

ggsave("SP Pre-Election Polling Outcomes Compared to the Election Result (2019).png", plot_all_sp_2019, width = 6, height = 4, dpi = 200, units = "in")


```

```{r}
#plot 2023
plot_all_sp_2023 <- ggplot() +
  geom_point(data = sp_polls_2023,
             aes(x = polling_period, y = poll_vote_share, shape = factor(polling_company)),
             size = 2) +
  geom_hline(yintercept = sp_result_2023, color = "red2", linewidth = 1) +
  geom_hline(yintercept = sp_pred_2023$mean_pred, linetype = "dashed", linewidth = 0.4) +
  geom_rect(data = sp_pred_2023,
            aes(xmin = -Inf, xmax = Inf, ymin = lower_ci *100, ymax = upper_ci *100),
            fill = "gray", alpha = 0.2, inherit.aes = FALSE) +
  geom_line(data = sp_ewma_2023,
            aes(x = polling_period, y = ewma),
            color = "black") +
  scale_y_continuous(limits = c(10, 35), breaks = seq(10, 35, by = 5), minor_breaks = seq(10, 35, by = 2.5), expand = expansion(0)) +
  labs(
    x = "Polling Period",
    y = "Vote Share (%)",
    shape = "Polling Company"
  ) +
  scale_shape_discrete(name = "", labels = company_labels) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    strip.background = element_rect(color = "white"),
    plot.margin = margin(t = 20, r = 20, b = 0, l = 10, unit = "pt")
  )


print(plot_all_sp_2023)

ggsave("SP Pre-Election Polling Outcomes Compared to the Election Result (2023).png", plot_all_sp_2023, width = 6, height = 4, dpi = 200, units = "in")


```

###SVP

```{r}
#plot 2019
plot_all_svp_2019 <- ggplot() +
  geom_point(data = svp_polls_2019,
             aes(x = polling_period, y = poll_vote_share, shape = factor(polling_company)),
             size = 2) +
  geom_hline(yintercept = svp_result_2019, color = "green4", linewidth = 1) +
  geom_hline(yintercept = svp_pred_2019$mean_pred, linetype = "dashed", linewidth = 0.4) +
  geom_rect(data = svp_pred_2019,
            aes(xmin = -Inf, xmax = Inf, ymin = lower_ci *100, ymax = upper_ci *100),
            fill = "gray", alpha = 0.2, inherit.aes = FALSE) +
  geom_line(data = svp_ewma_2019,
            aes(x = polling_period, y = ewma),
            color = "black") +
  scale_y_continuous(limits = c(15, 40), breaks = seq(15, 40, by = 5), minor_breaks = seq(15, 40, by = 2.5), expand = expansion(0)) +
  labs(
    x = "Polling Period",
    y = "Vote Share (%)",
    shape = "Polling Company"
  ) +
  scale_shape_discrete(name = "", labels = company_labels) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    strip.background = element_rect(color = "white"),
    plot.margin = margin(t = 20, r = 20, b = 0, l = 10, unit = "pt")
  )


print(plot_all_svp_2019)

ggsave("SVP Pre-Election Polling Outcomes Compared to the Election Result (2019).png", plot_all_svp_2019, width = 6, height = 4, dpi = 200, units = "in")


```

```{r}
#plot 2023
plot_all_svp_2023 <- ggplot() +
  geom_point(data = svp_polls_2023,
             aes(x = polling_period, y = poll_vote_share, shape = factor(polling_company)),
             size = 2) +
  geom_hline(yintercept = svp_result_2023, color = "green4", linewidth = 1) +
  geom_hline(yintercept = svp_pred_2023$mean_pred, linetype = "dashed", linewidth = 0.4) +
  geom_rect(data = svp_pred_2023,
            aes(xmin = -Inf, xmax = Inf, ymin = lower_ci *100, ymax = upper_ci *100),
            fill = "gray", alpha = 0.2, inherit.aes = FALSE) +
  geom_line(data = svp_ewma_2023,
            aes(x = polling_period, y = ewma),
            color = "black") +
  scale_y_continuous(limits = c(15, 40), breaks = seq(15, 40, by = 5), minor_breaks = seq(15, 40, by = 2.5), expand = expansion(0)) +
  labs(
    x = "Polling Period",
    y = "Vote Share (%)",
    shape = "Polling Company"
  ) +
  scale_shape_discrete(name = "", labels = company_labels) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    strip.background = element_rect(color = "white"),
    plot.margin = margin(t = 20, r = 20, b = 0, l = 10, unit = "pt")
  )


print(plot_all_svp_2023)

ggsave("SVP Pre-Election Polling Outcomes Compared to the Election Result (2023).png", plot_all_svp_2023, width = 6, height = 4, dpi = 200, units = "in")


```

###Other

```{r}
#plot 2019
plot_all_other_2019 <- ggplot() +
  geom_point(data = other_polls_2019,
             aes(x = polling_period, y = poll_vote_share, shape = factor(polling_company)),
             size = 2) +
  geom_hline(yintercept = other_result_2019, color = "plum2", linewidth = 1) +
  geom_hline(yintercept = other_pred_2019$mean_pred, linetype = "dashed", linewidth = 0.4) +
  geom_rect(data = other_pred_2019,
            aes(xmin = -Inf, xmax = Inf, ymin = lower_ci *100, ymax = upper_ci *100),
            fill = "gray", alpha = 0.2, inherit.aes = FALSE) +
  geom_line(data = other_ewma_2019,
            aes(x = polling_period, y = ewma),
            color = "black") +
  scale_y_continuous(limits = c(15, 40), breaks = seq(15, 40, by = 5), minor_breaks = seq(15, 40, by = 2.5), expand = expansion(0)) +
  labs(
    x = "Polling Period",
    y = "Vote Share (%)",
    shape = "Polling Company"
  ) +
  scale_shape_discrete(name = "", labels = company_labels) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    strip.background = element_rect(color = "white"),
    plot.margin = margin(t = 20, r = 20, b = 0, l = 10, unit = "pt")
  )


print(plot_all_other_2019)

ggsave("Other Pre-Election Polling Outcomes Compared to the Election Result (2019).png", plot_all_other_2019, width = 6, height = 4, dpi = 200, units = "in")


```

```{r}
#plot 2023
plot_all_other_2023 <- ggplot() +
  geom_point(data = other_polls_2023,
             aes(x = polling_period, y = poll_vote_share, shape = factor(polling_company)),
             size = 2) +
  geom_hline(yintercept = other_result_2023, color = "plum2", linewidth = 1) +
  geom_hline(yintercept = other_pred_2023$mean_pred, linetype = "dashed", linewidth = 0.4) +
  geom_rect(data = other_pred_2023,
            aes(xmin = -Inf, xmax = Inf, ymin = lower_ci *100, ymax = upper_ci *100),
            fill = "gray", alpha = 0.2, inherit.aes = FALSE) +
  geom_line(data = other_ewma_2023,
            aes(x = polling_period, y = ewma),
            color = "black") +
  scale_y_continuous(limits = c(10, 35), breaks = seq(10, 35, by = 5), minor_breaks = seq(10, 35, by = 2.5), expand = expansion(0)) +
  labs(
    x = "Polling Period",
    y = "Vote Share (%)",
    shape = "Polling Company"
  ) +
  scale_shape_discrete(name = "", labels = company_labels) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    strip.background = element_rect(color = "white"),
    plot.margin = margin(t = 20, r = 20, b = 0, l = 10, unit = "pt")
  )


print(plot_all_other_2023)

ggsave("Other Pre-Election Polling Outcomes Compared to the Election Result (2023).png", plot_all_other_2023, width = 6, height = 4, dpi = 200, units = "in")


```
